---
title: "Two_States_Two_Periods"
output: html_document
---


```{r}


```


Two states, treated state and control state. $s =0$ for control and $s=1$ for treated. Each state has $N$ counties.
Two periods, $t=0$ not treated and $t=1$ treated.

Here is the model for $y_{jst}$. Basically, there are three key ingredients in this model: two random effects(county random effects $\gamma_j$ and time random effects $\delta$), group-level predictors $\bar{X}_{s[j]}$, and treatment effects.


The group-level predictors capture the group varying intercept and the county random effects $\gamma_j$ accounts for the individual heterogenity not explained by the group-level predictors. $\lambda$ represents the time trend we need to model. In this two-period example, $\lambda$ is just a number. In a longer panel, $\lambda_t$ is a series of numbers could be modeled by iid normal.

$$
y_{jst}= A_0+ A_1^\top \bar{X}_{s[j]} + \gamma_j + \lambda I_{t=1} + \delta_j Z_{jst} + \epsilon_{jst} 
$$
Put usual normal prior on the regression coefficients $A_1$ and intercept $A_0$. Hierarchical normal prior on $\gamma_j$ and $\gamma_j$ has different mean and variance for different groups. Normal prior on $\delta_j$. Normal prior on $\lambda$. 


$\epsilon_{jst}$ are all iid normal(0,$\sigma_{y}^2$). 


Implement five different estimates:

1. Event Study. Take difference between the mean of treated states from time $1$ and time $0$.


$\bar{y}_{s=1,t=1} - \bar{y}_{s=1, t=0} = \bar\delta + \lambda + \bar{\epsilon}_{s=1,t=1} - \bar{\epsilon}_{s=1,t=0}\sim\mathcal{N}(\bar{\delta} + \lambda, \dfrac{2\sigma_{y}^2}{N}) $

2. Difference-in-mean. Take difference between the mean of treated states from time $1$ and mean of control states from time $1$

$$
\begin{equation}
\begin{aligned}
&\bar{y}_{s=1,t=1} - \bar{y}_{s=0, t=1} 
\\ = &  \bar\delta + A_1^\top (\bar{X}_{s=1} - \bar{X}_{s=0}) + \bar{\gamma}_{s=1} - \bar{\gamma}_{s=0} + \bar{\epsilon}_{s=1,t=1} - \bar{\epsilon}_{s=0,t=1}\\
\sim & \mathcal{N}(\bar\delta + A_1^\top (\bar{X}_{s=1} - \bar{X}_{s=0}) + \bar{\gamma}_{s=1} - \bar{\gamma}_{s=0}\;,\; \dfrac{2\sigma_{y}^2}{N}) 
\end{aligned}
\end{equation}
$$


3. DiD

$$
\begin{equation}
\begin{aligned}
& (\bar{y}_{s=1,t=1} - \bar{y}_{s=1, t=0}) - (\bar{y}_{s=0,t=1} - \bar{y}_{s=0, t=0}) \\
& = \bar\delta  + \bar{\epsilon}_{s=1,t=1} - \bar{\epsilon}_{s=1,t=0} - (\bar{\epsilon}_{s=0,t=1} - \bar{\epsilon}_{s=0,t=0})\\
& \sim\mathcal{N}(\bar{\delta} , \dfrac{4\sigma_{y}^2}{N}) 
\end{aligned}
\end{equation}
$$


4. Regression DiD controlling for something need to be figured out.



5. Bayesian DiD.

We expect that conditional on the hyperparameters the Bayesian DiD estimate should be of form of the something subtracted by weighted $\lambda_t$(time trend) and $\bar{\gamma}_{s=1} - \bar{\gamma}_{s=0}$ and $ A_1^\top (\bar{X}_{s=1} - \bar{X}_{s=0})$(the group level variation).



```{r}


# =========================
# Libraries
# =========================
library(cmdstanr)
library(posterior)
library(dplyr)
library(tidyr)
library(purrr)
library(ggplot2)

# =========================
# Compile Stan once
# =========================
stan_file <- "Two_States_Two_Periods.stan"
mod <- cmdstan_model(stan_file)

# =========================
# Simulation (2x2 with Xbar in gamma-mean and u_s RE)
# =========================
simulate_2x2_with_x <- function(N_per_state = 40,
                                K = 2,
                                Xbar = NULL,      # S x K matrix (S=2). If NULL, draw N(0,1)
                                A0 = 0.0,
                                lambda = 0.5,
                                mu_delta = 1.0,
                                tau_delta = 0.6,
                                alpha_gamma = 0.0,
                                B = NULL,         # length K; if NULL, draw N(0,0.5)
                                sigma_u = 0.4,
                                sigma_gamma = 1.0,
                                sigma_y = 1.0,
                                seed = 1) {
  set.seed(seed)
  S <- 2L
  J_t <- N_per_state
  J_c <- N_per_state
  J   <- J_t + J_c
  is_treated <- c(rep(1L, J_t), rep(0L, J_c))
  state_of_county <- c(rep(1L, J_t), rep(2L, J_c))

  if (is.null(Xbar)) Xbar <- matrix(rnorm(S * K, 0, 1), nrow = S, ncol = K)
  if (is.null(B))    B    <- rnorm(K, 0, 0.5)
  u <- rnorm(S, 0, sigma_u)

  # gamma prior mean (by county), then center to match sum-to-zero identification
  mu_gamma <- alpha_gamma + Xbar[state_of_county, , drop = FALSE] %*% B + u[state_of_county]
  mu_gamma <- as.numeric(mu_gamma)
  mu_gamma_cen <- mu_gamma - mean(mu_gamma)

  # draw gamma_j around centered mean, then enforce exact sum-to-zero
  gamma <- rnorm(J, mean = mu_gamma_cen, sd = sigma_gamma)
  gamma <- gamma - mean(gamma)

  # deltas: treated counties ~ N(mu_delta, tau_delta^2), controls = 0
  delta <- rep(0, J)
  delta[is_treated == 1L] <- rnorm(J_t, mu_delta, tau_delta)

  # 2x2 panel rows
  county <- rep(seq_len(J), each = 2)
  post   <- rep(c(0L, 1L), times = J)
  D      <- is_treated[county] * post

  mu <- A0 + gamma[county] + lambda * post + delta[county] * D
  y  <- rnorm(2 * J, mu, sigma_y)

  list(
    stan_data = list(
      N = length(y), J = J, S = S, K = K,
      y = y, county = as.integer(county), post = as.integer(post), D = as.integer(D),
      state_of_county = as.integer(state_of_county),
      Xbar = Xbar,
      is_treated = as.integer(is_treated),
      J_treated = J_t,
      treated_idx = as.integer(which(is_treated == 1L))
    ),
    truth = list(
      A0 = A0, lambda = lambda, mu_delta = mu_delta, tau_delta = tau_delta,
      alpha_gamma = alpha_gamma, B = B, u = u, sigma_u = sigma_u,
      sigma_gamma = sigma_gamma, sigma_y = sigma_y,
      gamma = gamma, delta = delta, Xbar = Xbar,
      is_treated = is_treated
    )
  )
}

# =========================
# Classical estimators (ATT) and per-county d_j
# =========================
compute_classical <- function(y, county, post, state, sigma_y, N_per_state) {
  # cell means
  y_11 <- mean(y[state == 1 & post == 1])
  y_10 <- mean(y[state == 1 & post == 0])
  y_01 <- mean(y[state == 0 & post == 1])
  y_00 <- mean(y[state == 0 & post == 0])

  # ATT competitors
  ES_hat  <- y_11 - y_10
  DM_hat  <- y_11 - y_01
  DiD_hat <- (y_11 - y_10) - (y_01 - y_00)

  z <- qnorm(0.975)
  ES_se  <- sqrt(2 * sigma_y^2 / N_per_state)
  DM_se  <- sqrt(2 * sigma_y^2 / N_per_state)
  DiD_se <- sqrt(4 * sigma_y^2 / N_per_state)

  list(
    ES  = c(est = ES_hat,  lo = ES_hat - z * ES_se,  hi = ES_hat + z * ES_se),
    DM  = c(est = DM_hat,  lo = DM_hat - z * DM_se,  hi = DM_hat + z * DM_se),
    DiD = c(est = DiD_hat, lo = DiD_hat - z * DiD_se, hi = DiD_hat + z * DiD_se)
  )
}

# Per-county DiD contrasts for treated counties:
# d_j = (y_j11 - y_j10) - (mean_control_1 - mean_control_0)
compute_per_county_dj <- function(y, county, post, state, sigma_y, N_per_state, treated_ids) {
  y_c_1 <- mean(y[state == 0 & post == 1])
  y_c_0 <- mean(y[state == 0 & post == 0])

  est <- numeric(length(treated_ids))
  lo  <- hi <- est

  z <- qnorm(0.975)
  # Var(d_j) = Var(y_j11 - y_j10) + Var(mean_c1 - mean_c0) = 2 sigma_y^2 + 2 sigma_y^2 / N
  se_dj <- sqrt(2 * sigma_y^2 + 2 * sigma_y^2 / N_per_state)

  for (k in seq_along(treated_ids)) {
    j <- treated_ids[k]
    y_j1 <- y[county == j & post == 1]
    y_j0 <- y[county == j & post == 0]
    dj <- (y_j1 - y_j0) - (y_c_1 - y_c_0)
    est[k] <- dj
    lo[k]  <- dj - z * se_dj
    hi[k]  <- dj + z * se_dj
  }
  data.frame(county = treated_ids, est = est, lo = lo, hi = hi)
}

# =========================
# Fit Stan once per dataset, extract summaries
# =========================
fit_bayes_once <- function(stan_data,
                           iter_warmup = 750, iter_sampling = 750,
                           chains = 2, seed = 1234) {
  
  outdir <- file.path(getwd(), "cmdstan_out")
  dir.create(outdir, showWarnings = FALSE, recursive = TRUE)
  
  fit <- mod$sample(
    data = stan_data,
    chains = chains, parallel_chains = chains,
    iter_warmup = iter_warmup, iter_sampling = iter_sampling,
    refresh = 0, show_messages = FALSE,output_dir = outdir,
  save_cmdstan_config = TRUE,   # <- key
  save_warmup = TRUE,                  # ensure CSVs persist
    seed = seed
  )
  dr <- as_draws_df(fit$draws())

  # Scalars
  ATT_post    <- dr$ATT
  lambda_post <- dr$lambda

  # Vectors
  delta_cols <- grep("^delta\\[", names(dr), value = TRUE)
  gamma_cols <- grep("^gamma\\[", names(dr), value = TRUE)

  list(
    fit = fit,
    post = list(
      ATT    = list(mean = mean(ATT_post),
                    ci = quantile(ATT_post, c(0.025, 0.975))),
      lambda = list(mean = mean(lambda_post),
                    ci = quantile(lambda_post, c(0.025, 0.975))),
      delta  = list(mean = sapply(delta_cols, \(nm) mean(dr[[nm]])),
                    lo   = sapply(delta_cols, \(nm) quantile(dr[[nm]], 0.025)),
                    hi   = sapply(delta_cols, \(nm) quantile(dr[[nm]], 0.975))),
      gamma  = list(mean = sapply(gamma_cols, \(nm) mean(dr[[nm]])),
                    lo   = sapply(gamma_cols, \(nm) quantile(dr[[nm]], 0.025)),
                    hi   = sapply(gamma_cols, \(nm) quantile(dr[[nm]], 0.975)))
    )
  )
}

# =========================
# One replication wrapper
# =========================
one_replication <- function(N_per_state = 40, K = 2, seed = 1,
                            iter_warmup = 750, iter_sampling = 750, chains = 2) {
  sim <- simulate_2x2_with_x(N_per_state = N_per_state, K = K, seed = seed)

  # Classical
  county <- rep(seq_len(sim$stan_data$J), each = 2)
  post   <- sim$stan_data$post
  state  <- sim$stan_data$state_of_county - 1L
  y      <- sim$stan_data$y
  Nps    <- N_per_state
  sigy   <- sim$truth$sigma_y

  cls <- compute_classical(y, county, post, state, sigy, Nps)
  dj  <- compute_per_county_dj(y, county, post, state, sigy, Nps,
                               treated_ids = which(sim$truth$is_treated == 1L))

  # Bayes
  fitb <- fit_bayes_once(sim$stan_data,
                         iter_warmup = iter_warmup, iter_sampling = iter_sampling,
                         chains = chains, seed = 1000 + seed)

  # --- Metrics ---
  truth <- sim$truth

  # ATT metrics
  ATT_true <- mean(truth$delta[truth$is_treated == 1L])
  att_bayes <- fitb$post$ATT
  att_metrics <- tibble(
    method = c("ES","DM","DiD","Bayes"),
    est    = c(cls$ES["est"], cls$DM["est"], cls$DiD["est"], att_bayes$mean),
    lo     = c(cls$ES["lo"],  cls$DM["lo"],  cls$DiD["lo"],  att_bayes$ci[1]),
    hi     = c(cls$ES["hi"],  cls$DM["hi"],  cls$DiD["hi"],  att_bayes$ci[2])
  ) %>%
    mutate(true = ATT_true,
           bias = est - true,
           sqerr = (est - true)^2,
           covered = as.integer(true >= lo & true <= hi),
           width = hi - lo)

  # lambda metrics (only Bayes estimates lambda here)
  lam_true <- truth$lambda
  lam <- fitb$post$lambda
  lambda_metrics <- tibble(
    method = "Bayes", est = lam$mean, lo = lam$ci[1], hi = lam$ci[2],
    true = lam_true, bias = est - true, sqerr = (est - true)^2,
    covered = as.integer(true >= lo & true <= hi), width = hi - lo
  )

  # Per-county delta metrics (treated counties)
  treated_ids <- which(truth$is_treated == 1L)
  # Bayesian
  d_bayes <- tibble(
    county = treated_ids,
    est = fitb$post$delta$mean[treated_ids],
    lo  = fitb$post$delta$lo[treated_ids],
    hi  = fitb$post$delta$hi[treated_ids],
    true = truth$delta[treated_ids]
  ) %>%
    mutate(bias = est - true, sqerr = (est - true)^2,
           covered = as.integer(true >= lo & true <= hi),
           width = hi - lo,
           method = "Bayes")
  # Frequentist per-county (d_j)
  d_freq <- dj %>%
    mutate(true = truth$delta[county],
           bias = est - true, sqerr = (est - true)^2,
           covered = as.integer(true >= lo & true <= hi),
           width = hi - lo,
           method = "Per-county DiD")

  # County FE gamma metrics (Bayes only; compare to truth)
  g_bayes <- tibble(
    county = seq_along(truth$gamma),
    est = fitb$post$gamma$mean,
    lo  = fitb$post$gamma$lo,
    hi  = fitb$post$gamma$hi,
    true = truth$gamma
  ) %>%
    mutate(bias = est - true, sqerr = (est - true)^2,
           covered = as.integer(true >= lo & true <= hi),
           width = hi - lo,
           method = "Bayes")

  list(att = att_metrics,
       lambda = lambda_metrics,
       delta = bind_rows(d_bayes, d_freq),
       gamma = g_bayes)
}

```


```{r}

# =========================
# Run R replications and summarize
# =========================
run_simulation <- function(R = 100, N_per_state = 40, K = 2,
                           iter_warmup = 500, iter_sampling = 500, chains = 2,
                           seed0 = 123) {
  res <- vector("list", R)
  for (r in seq_len(R)) {
    res[[r]] <- one_replication(
      N_per_state = N_per_state, K = K, seed = seed0 + r,
      iter_warmup = iter_warmup, iter_sampling = iter_sampling, chains = chains
    )
    if (r %% 5 == 0) cat("Completed", r, "of", R, "\n")
  }
  # Bind and summarize
  ATT_all    <- bind_rows(lapply(res, `[[`, "att"), .id = "rep")
  LAMBDA_all <- bind_rows(lapply(res, `[[`, "lambda"), .id = "rep")
  DELTA_all  <- bind_rows(lapply(res, `[[`, "delta"), .id = "rep")
  GAMMA_all  <- bind_rows(lapply(res, `[[`, "gamma"), .id = "rep")

  # ATT summary across reps
  att_sum <- ATT_all %>%
    group_by(method) %>%
    summarize(
      mean_est = mean(est),
      bias = mean(bias),
      rmse = sqrt(mean(sqerr)),
      cover95 = mean(covered),
      avg_width = mean(width),
      .groups = "drop"
    )

  # lambda summary (Bayes only)
  lambda_sum <- LAMBDA_all %>%
    summarize(
      mean_est = mean(est),
      bias = mean(bias),
      rmse = sqrt(mean(sqerr)),
      cover95 = mean(covered),
      avg_width = mean(width)
    )

  # delta summary (average across treated counties and reps)
  delta_sum <- DELTA_all %>%
    group_by(method) %>%
    summarize(
      mean_bias = mean(bias),
      rmse = sqrt(mean(sqerr)),
      cover95 = mean(covered),
      avg_width = mean(width),
      .groups = "drop"
    )

  # gamma summary (Bayes only; averaged over counties and reps)
  gamma_sum <- GAMMA_all %>%
    summarize(
      mean_bias = mean(bias),
      rmse = sqrt(mean(sqerr)),
      cover95 = mean(covered),
      avg_width = mean(width)
    )

  list(ATT = att_sum, LAMBDA = lambda_sum, DELTA = delta_sum, GAMMA = gamma_sum,
       raw = list(ATT = ATT_all, LAMBDA = LAMBDA_all, DELTA = DELTA_all, GAMMA = GAMMA_all))
}
```



```{r}
# =========================
# Optional quick plots
# =========================
plot_att_summary <- function(att_sum) {
  ggplot(att_sum, aes(x = method, y = rmse, fill = method)) +
    geom_col() +
    geom_text(aes(label = sprintf("bias=%.3f\ncov=%.2f\nw=%.3f", bias, cover95, avg_width)),
              vjust = -0.3, size = 3) +
    labs(title = "ATT: RMSE (labels show bias, coverage, CI width)", x = NULL, y = "RMSE") +
    theme_minimal(base_size = 12) + theme(legend.position = "none")
}

plot_delta_summary <- function(delta_sum) {
  ggplot(delta_sum, aes(x = method, y = rmse, fill = method)) +
    geom_col() +
    geom_text(aes(label = sprintf("bias=%.3f\ncov=%.2f\nw=%.3f", mean_bias, cover95, avg_width)),
              vjust = -0.3, size = 3) +
    labs(title = expression("Per-county "~delta[j]~": RMSE (bias, coverage, width)"),
         x = NULL, y = "RMSE") +
    theme_minimal(base_size = 12) + theme(legend.position = "none")
}

```


```{r}

# =========================
# Example: run 100 reps (adjust MCMC settings as needed)
# =========================
# NOTE: 100 MCMC fits can be time-consuming. Start small to smoke test:
sim_out <- run_simulation(R = 100, N_per_state = 40, K = 2, iter_warmup = 500, iter_sampling = 2000, chains = 2)
# Then scale up:
# sim_out <- run_simulation(R = 100, N_per_state = 40, K = 2, iter_warmup = 500, iter_sampling = 500, chains = 2)
```








$$
y_{jst}= A_0+ A_1^\top \bar{X}_{s[j]} + A_2^\top X_j + \gamma_j + \lambda_t + \delta_j Z_{jst} + \epsilon_{jst} 
$$




# New trunk of code

```{r}
# 2×2: one treated state (s=1), one control (s=0)
simulate_2x2_grouped <- function(N_per_state=40, K=1,
                                 A0=0.0, A1=0.4,
                                 Xbar = matrix(c(-0.5, 0.8), nrow=2, byrow=TRUE),
                                 # state RE scales (truth)
                                 sigma_u_treat=2, sigma_u_ctrl=1.4,
                                 # county RE scales by state (truth)
                                 sigma_c_treat=1.0, sigma_c_ctrl=0.6,
                                 # treatment & time
                                 lambda=0.5, mu_delta=3.0, sigma_delta=1.6,
                                 sigma_y=1.0, seed=1L){
  set.seed(seed)
  S <- 2L; J_t <- N_per_state; J_c <- N_per_state; J <- J_t + J_c
  is_treated <- c(rep(1L,J_t), rep(0L,J_c))
  state_of_county <- c(rep(1L,J_t), rep(2L,J_c))
  state_is_treated <- c(1L,0L)

  # truth: u_s and c_j by state group
  u <- c(rnorm(1,0,sigma_u_treat), rnorm(1,0,sigma_u_ctrl))
  c <- numeric(J)
  c[is_treated==1L] <- rnorm(J_t,0,sigma_c_treat)
  c[is_treated==0L] <- rnorm(J_c,0,sigma_c_ctrl)

  # deltas (treated counties only)
  delta <- rep(0,J)
  delta[is_treated==1L] <- rnorm(J_t, mu_delta, sigma_delta)

  # panel rows
  county <- rep(seq_len(J), each=2)
  post   <- rep(c(0L,1L), times=J)
  D      <- is_treated[county]*post
  s_row  <- state_of_county[county]

  A1_vec <- if (length(A1)==1) rep(A1,K) else as.numeric(A1)
  lin_x  <- rowSums(Xbar[s_row,,drop=FALSE] *
                    matrix(A1_vec, nrow=length(s_row), ncol=K, byrow=TRUE))

  mu <- A0 + lin_x + u[s_row] + c[county] + lambda*post + delta[county]*D
  y  <- rnorm(length(mu), mu, sigma_y)

  list(
    stan_data = list(
      N=length(y), J=J, S=2L, K=K, y=y,
      county=as.integer(county), post=as.integer(post), D=as.integer(D),
      state_of_county=as.integer(state_of_county),
      Xbar=Xbar,
      is_treated=as.integer(is_treated),
      state_is_treated=as.integer(state_is_treated),
      J_treated=J_t, treated_idx=as.integer(which(is_treated==1L)),
      center_c=0L,
      pool_u_within_group=0L,     # 2×2: no pooling across groups
      big_sd_u=1e3                # diffuse prior per state
    ),
    truth = list(
      A0=A0, A1=A1_vec, Xbar=Xbar, lambda=lambda, sigma_y=sigma_y,
      u=u, c=c, is_treated=is_treated, delta=delta,
      mu_delta=mu_delta, sigma_delta=sigma_delta
    )
  )
}

```



```{r}
library(cmdstanr)
library(posterior)
library(dplyr)
library(tidyr)
library(purrr)

# Classical ATT estimators (+ Wald CIs under homoskedastic σ_y)
att_classical <- function(y, county, post, state, sigma_y, N_per_state){
  y_11 <- mean(y[state==1 & post==1]); y_10 <- mean(y[state==1 & post==0])
  y_01 <- mean(y[state==0 & post==1]); y_00 <- mean(y[state==0 & post==0])
  ES  <- y_11 - y_10
  DM  <- y_11 - y_01
  DiD <- (y_11 - y_10) - (y_01 - y_00)
  z <- qnorm(0.975)
  ES_se  <- sqrt(2 * sigma_y^2 / N_per_state)
  DM_se  <- sqrt(2 * sigma_y^2 / N_per_state)
  DiD_se <- sqrt(4 * sigma_y^2 / N_per_state)
  tibble(
    method = c("ES","DM","DiD"),
    est    = c(ES, DM, DiD),
    lo     = c(ES - z*ES_se,  DM - z*DM_se,  DiD - z*DiD_se),
    hi     = c(ES + z*ES_se,  DM + z*DM_se,  DiD + z*DiD_se)
  )
}

# Per-county DiD for treated counties (analytic CI under σ_y)
per_county_did <- function(y, county, post, state, sigma_y, N_per_state, treated_ids){
  y_c1 <- mean(y[state==0 & post==1]); y_c0 <- mean(y[state==0 & post==0])
  z <- qnorm(0.975)
  se_dj <- sqrt(2*sigma_y^2 + 2*sigma_y^2/N_per_state)
  map_dfr(treated_ids, function(j){
    y_j1 <- y[county==j & post==1]; y_j0 <- y[county==j & post==0]
    dj <- (y_j1 - y_j0) - (y_c1 - y_c0)
    tibble(county=j, est=dj, lo=dj - z*se_dj, hi=dj + z*se_dj)
  })
}

```



```{r}

# ---- one-time: pick a durable folder for chain CSVs ----
.default_outdir <- file.path(getwd(), "cmdstan_out")
if (!dir.exists(.default_outdir)) dir.create(.default_outdir, recursive = TRUE)

fit_bayes_once <- function(stan_data,
                           treated_idx  = NULL,      # will try stan_data$treated_idx if NULL
                           delta_name   = "delta",
                           iter_warmup  = 500,
                           iter_sampling= 1000,
                           chains       = 4,
                           parallel_chains = chains,
                           seed         = 2025,
                           output_dir   = .default_outdir,
                           refresh      = 0) {

  # --- resolve treated indices ---
  if (is.null(treated_idx)) {
    treated_idx <- stan_data$treated_idx %||% stan_data$treated_ids
    if (is.null(treated_idx)) stop("Provide treated_idx or include $treated_idx in stan_data.")
  }
  treated_idx <- as.integer(treated_idx)
  J1 <- length(treated_idx)

  # --- unique chain ids to avoid filename collisions ---
  chain_ids <- seq_len(chains)

  # --- run sampler with pinned output_dir ---
  
  outdir <- file.path(getwd(), "cmdstan_out")
  dir.create(outdir, showWarnings = FALSE, recursive = TRUE)
  
  fit <- mod$sample(
    data = stan_data,
    chains = chains, parallel_chains = chains,
    iter_warmup = iter_warmup, iter_sampling = iter_sampling,
    refresh = 0, show_messages = FALSE,output_dir = outdir,
  save_cmdstan_config = TRUE,   # <- key
  save_warmup = TRUE,           # optional
  output_basename = "run1",    # optional: stable filenames
  seed = seed
  )

  # --- check CSVs exist & chains returned 0 ---
  csvs <- fit$output_files(include_failed = TRUE)
  if (!all(file.exists(csvs))) {
    message("Some chain CSVs are missing:\n", paste(csvs[!file.exists(csvs)], collapse = "\n"))
  }
  rc <- fit$return_codes()
  if (any(rc != 0L) || !all(file.exists(csvs))) {
    # print stderr/stdout paths to help debug
    std_out <- fit$stdout_files()
    diag_files <- fit$diagnostic_files()
    message("Nonzero return codes: ", paste(rc, collapse = ", "))
    message("Stdout files:\n", paste(std_out, collapse = "\n"))
    message("Diagnostic files:\n", paste(diag_files, collapse = "\n"))
    # Try a safer rerun (single chain) to surface the underlying error
    message("Retrying with chains=1, parallel_chains=1, refresh=0 to surface messages...")
    fit <- mod$sample(
      data = stan_data,
      seed = seed + 1L,
      chains = 1,
      parallel_chains = 1,
      chain_ids = 1,
      iter_warmup = iter_warmup,
      iter_sampling = iter_sampling,
      refresh = 0,
      output_dir = output_dir
    )
  }

  # --- extract delta draws ONLY (avoid delta_raw) ---
  dr_delta <- posterior::as_draws_df(fit$draws(variables = delta_name))
  delta_cols <- grep(paste0("^", delta_name, "\\[[0-9]+\\]$"),
                     names(dr_delta), value = TRUE)
  if (length(delta_cols) == 0) {
    stop("No columns found for parameter '", delta_name, "'. Check Stan variable names.")
  }

  j_stan <- as.integer(sub(paste0("^", delta_name, "\\[([0-9]+)\\]$"), "\\1", delta_cols))
  ord <- order(j_stan)
  delta_cols <- delta_cols[ord]; j_stan <- j_stan[ord]
  Delta_mat <- as.matrix(dr_delta[, delta_cols, drop = FALSE])

  # --- handle two cases: delta over ALL counties vs TREATED-only ---
  if (!is.null(stan_data$J) && length(j_stan) == stan_data$J) {
    pos_treat <- match(treated_idx, j_stan)
    if (anyNA(pos_treat)) {
      stop("Some treated_idx are not present among indices of '", delta_name, "'.")
    }
    Delta_tr <- Delta_mat[, pos_treat, drop = FALSE]
    county_labels <- treated_idx
  } else if (length(j_stan) == J1 && all(j_stan == seq_len(J1))) {
    # Stan stored delta only for treated counties
    Delta_tr <- Delta_mat
    county_labels <- treated_idx
  } else {
    stop("Dimension mismatch: ", delta_name, " has ", length(j_stan),
         " entries; expected J (all) or J1 (treated).")
  }

  # --- per-county posteriors (treated only) ---
  delta_mean <- colMeans(Delta_tr)
  delta_sd   <- apply(Delta_tr, 2, stats::sd)
  delta_l95  <- apply(Delta_tr, 2, stats::quantile, probs = 0.025)
  delta_u95  <- apply(Delta_tr, 2, stats::quantile, probs = 0.975)

  delta_tbl <- tibble::tibble(
    j   = as.integer(county_labels),
    est = as.numeric(delta_mean),
    sd  = as.numeric(delta_sd),
    lo  = as.numeric(delta_l95),
    hi  = as.numeric(delta_u95)
  )

  # --- ATT from the same draws (coherent) ---
  ATT_draws <- rowMeans(Delta_tr)
  ATT_mean  <- mean(ATT_draws)
  ATT_ci    <- stats::quantile(ATT_draws, c(0.025, 0.975), names = FALSE)

  list(
    ATT = list(mean = ATT_mean, lo = ATT_ci[1], hi = ATT_ci[2], draws = ATT_draws),
    delta = delta_tbl,
    output_files = fit$output_files()
  )
}

```



```{r}
quick_compare <- function(R = 50, N_per_state = 40, K = 1,
                          A0 = 0.0, A1 = 0.4,
                          Xbar = matrix(c(-0.5, 0.8), nrow = 2, byrow = TRUE),
                          sigma_u_treat = 1.4, sigma_u_ctrl = 0.4,
                          sigma_c_treat = 1.0, sigma_c_ctrl = 0.7,
                          lambda = 2, mu_delta = 1.0, sigma_delta = 0.6,
                          sigma_y = 1.0, seed0 = 1234,
                          iter_warmup = 500, iter_sampling = 500, chains = 2) {

  ATT_rows   <- vector("list", R)
  DELTA_rows <- vector("list", R)

  for (r in seq_len(R)) {
    sim <- simulate_2x2_grouped(N_per_state, K, A0, A1, Xbar,
                                sigma_u_treat, sigma_u_ctrl,
                                sigma_c_treat, sigma_c_ctrl,
                                lambda, mu_delta, sigma_delta, sigma_y,
                                seed = seed0 + r)

    ## classical (ATT + per-county DiD)
    county <- rep(seq_len(sim$stan_data$J), each = 2)
    post   <- sim$stan_data$post
    state  <- sim$stan_data$state_of_county - 1L
    y      <- sim$stan_data$y
    Nps    <- N_per_state
    treated_ids <- which(sim$truth$is_treated == 1L)

    cls  <- att_classical(y, county, post, state, sigma_y, Nps)                    # data.frame with {method, est, lo, hi}
    dj   <- per_county_did(y, county, post, state, sigma_y, Nps, treated_ids)      # tibble {county, est, lo, hi}

    ## Bayes (per-county posteriors + ATT from the same draws)
    fitb <- fit_bayes_once(
      stan_data       = sim$stan_data,
      treated_idx     = sim$stan_data$treated_idx,
      iter_warmup     = iter_warmup,
      iter_sampling   = iter_sampling,
      chains          = chains,
      parallel_chains = chains,
      seed            = seed0 + 10000 + r,
      refresh         = 0
    )

    ## --- ATT metrics ---
    true_ATT <- mean(sim$truth$delta[treated_ids])
    att_all <- dplyr::bind_rows(
      cls,
      tibble::tibble(method = "Bayes", est = fitb$ATT$mean, lo = fitb$ATT$lo, hi = fitb$ATT$hi)
    ) |>
      dplyr::mutate(true = true_ATT,
                    bias = est - true,
                    sqerr = (est - true)^2,
                    covered = as.integer(true >= lo & true <= hi),
                    width = hi - lo)
    ATT_rows[[r]] <- att_all

    ## --- δ_j metrics (treated counties): DiD vs Bayes ---
    db <- fitb$delta |>
      dplyr::filter(j %in% treated_ids) |>
      dplyr::rename(est_b = est, lo_b = lo, hi_b = hi)
    df <- dj |>
      dplyr::rename(j = county, est_f = est, lo_f = lo, hi_f = hi)

    d_join <- dplyr::inner_join(db, df, by = "j") |>
      dplyr::mutate(
        truth = sim$truth$delta[j],
        ## significance (two-sided 5%)
        sig_b = (lo_b > 0) | (hi_b < 0),
        sig_f = (lo_f > 0) | (hi_f < 0),
        ## Type-S indicators among the significant
        typeS_b = sig_b & (sign(est_b) != sign(truth)),
        typeS_f = sig_f & (sign(est_f) != sign(truth)),
        ## squared errors
        sq_b = (est_b - truth)^2,
        sq_f = (est_f - truth)^2
      )

    n_sig_b <- sum(d_join$sig_b)
    n_sig_f <- sum(d_join$sig_f)
    typeS_cnt_b <- sum(d_join$typeS_b)
    typeS_cnt_f <- sum(d_join$typeS_f)

    DELTA_rows[[r]] <- tibble::tibble(
      ## Counts
      n_sig_b = n_sig_b,
      n_sig_f = n_sig_f,
      ## Type-S counts and rates (NA if no significant)
      typeS_b_count = typeS_cnt_b,
      typeS_f_count = typeS_cnt_f,
      typeS_b_rate  = ifelse(n_sig_b > 0, typeS_cnt_b / n_sig_b, NA_real_),
      typeS_f_rate  = ifelse(n_sig_f > 0, typeS_cnt_f / n_sig_f, NA_real_),
      ## MSE and RMSE
      mse_b   = mean(d_join$sq_b),
      mse_f   = mean(d_join$sq_f),
      rmse_b  = sqrt(mean(d_join$sq_b)),
      rmse_f  = sqrt(mean(d_join$sq_f)),
      ## Coverage and average width (kept for completeness)
      cover_b = mean(d_join$truth >= db$lo_b & d_join$truth <= db$hi_b),
      cover_f = mean(d_join$truth >= df$lo_f & d_join$truth <= df$hi_f),
      width_b = mean(db$hi_b - db$lo_b),
      width_f = mean(df$hi_f - df$lo_f),
      ## Spearman rank correlations
      rank_b  = suppressWarnings(cor(d_join$est_b, d_join$truth, method = "spearman")),
      rank_f  = suppressWarnings(cor(d_join$est_f, d_join$truth, method = "spearman"))
    )
  }

  ATT_all   <- dplyr::bind_rows(ATT_rows)
  DELTA_all <- dplyr::bind_rows(DELTA_rows)

  att_sum <- ATT_all |>
    dplyr::group_by(method) |>
    dplyr::summarize(
      rmse = sqrt(mean(sqerr)),
      mse  = mean(sqerr),
      bias = mean(bias),
      cover95 = mean(covered),
      avg_width = mean(width),
      .groups = "drop"
    )

  delta_sum <- DELTA_all |>
    dplyr::summarize(
      ## Bayes
      n_sig_b_mean = mean(n_sig_b),
      typeS_b_count_mean = mean(typeS_b_count),
      typeS_b_rate_mean  = mean(typeS_b_rate, na.rm = TRUE),
      mse_b = mean(mse_b), rmse_b = mean(rmse_b),
      cover_b = mean(cover_b), width_b = mean(width_b),
      rank_b  = mean(rank_b, na.rm = TRUE),
      ## DiD
      n_sig_f_mean = mean(n_sig_f),
      typeS_f_count_mean = mean(typeS_f_count),
      typeS_f_rate_mean  = mean(typeS_f_rate, na.rm = TRUE),
      mse_f = mean(mse_f), rmse_f = mean(rmse_f),
      cover_f = mean(cover_f), width_f = mean(width_f),
      rank_f  = mean(rank_f, na.rm = TRUE)
    )

  list(ATT = att_sum,
       DELTA = delta_sum,
       ATT_raw = ATT_all,
       DELTA_raw = DELTA_all)
}


```



```{r}
library(purrr); library(dplyr); library(tidyr)

run_grid <- function(lambdas, A_grid, R = 50, N_per_state = 40, ...) {
  map2_dfr(lambdas, A_grid, \(lam, a1) {
    res <- quick_compare(R = R, N_per_state = N_per_state, A1 = a1, lambda = lam, ...)
    # Flatten: keep only delta summaries for plotting the 4 quantities
    tibble(lambda = lam, A1 = a1) |>
      bind_cols(res$DELTA)
  })
}

# Example:
# lambdas <- seq(0, 2, by = 0.25)
# A_grid  <- rep(0.4, length(lambdas))   # hold A1 fixed, or vary if you want
# grid_out <- run_grid(lambdas, A_grid, R = 50, iter_warmup = 300, iter_sampling = 600, chains = 2)



```


```{r}

library(ggplot2)

plot_delta_metrics <- function(grid_out, facet_by = NULL, use_typeS = c("count","rate")) {
  use_typeS <- match.arg(use_typeS)
  df <- grid_out

  # rename columns to a common long form for plotting
  long_list <- list(
    n_sig   = df |> select(lambda, {{facet_by}}, n_sig_b_mean, n_sig_f_mean) |>
      pivot_longer(cols = c(n_sig_b_mean, n_sig_f_mean),
                   names_to = "method", values_to = "value") |>
      mutate(metric = "Number significant",
             method = ifelse(method == "n_sig_b_mean", "Bayes", "DiD")),
    typeS   = {
      if (use_typeS == "count") {
        df |> select(lambda, {{facet_by}}, typeS_b_count_mean, typeS_f_count_mean) |>
          pivot_longer(cols = c(typeS_b_count_mean, typeS_f_count_mean),
                       names_to = "method", values_to = "value") |>
          mutate(metric = "Type-S (count)",
                 method = ifelse(method == "typeS_b_count_mean", "Bayes", "DiD"))
      } else {
        df |> select(lambda, {{facet_by}}, typeS_b_rate_mean, typeS_f_rate_mean) |>
          pivot_longer(cols = c(typeS_b_rate_mean, typeS_f_rate_mean),
                       names_to = "method", values_to = "value") |>
          mutate(metric = "Type-S (rate)",
                 method = ifelse(method == "typeS_b_rate_mean", "Bayes", "DiD"))
      }
    },
    mse     = df |> select(lambda, {{facet_by}}, mse_b, mse_f) |>
      pivot_longer(cols = c(mse_b, mse_f),
                   names_to = "method", values_to = "value") |>
      mutate(metric = "MSE", method = ifelse(method == "mse_b", "Bayes", "DiD")),
    rank    = df |> select(lambda, {{facet_by}}, rank_b, rank_f) |>
      pivot_longer(cols = c(rank_b, rank_f),
                   names_to = "method", values_to = "value") |>
      mutate(metric = "Spearman ρ", method = ifelse(method == "rank_b", "Bayes", "DiD"))
  )

  plot_df <- bind_rows(long_list)

  p <- ggplot(plot_df, aes(x = lambda, y = value, linetype = method, shape = method)) +
    geom_line(linewidth = 0.9) +
    geom_point(size = 1.8) +
    scale_linetype_manual(values = c(Bayes = "dashed", DiD = "solid")) +
    scale_shape_manual(values = c(Bayes = 17, DiD = 16)) +
    labs(x = expression(lambda), y = NULL, linetype = "Method", shape = "Method") +
    theme_minimal(base_size = 12) +
    theme(legend.position = "bottom") +
    facet_wrap(~ metric, scales = "free_y", ncol = 2)

  # Optional facet by A1/A2 if it varies
  if (!is.null(facet_by)) {
    p <- p + facet_grid(rows = vars(metric), cols = vars(!!rlang::sym(facet_by)),
                        scales = "free_y")
  }
  p
}
```

```{r}
# Example usage:

#Example:
lambdas <- seq(0, 2, by = 0.25)
A_grid  <- rep(0.4, length(lambdas))   # hold A1 fixed, or vary if you want
grid_out <- run_grid(lambdas, A_grid, R = 1, iter_warmup = 300, iter_sampling = 600, chains = 2)
p <- plot_delta_metrics(grid_out, facet_by = "A1", use_typeS = "rate")
print(p)

```

produce the same four plots as the chicken paper.

About ATT.


Five competitors:

+ Event study with standard error
+ Difference-in-means with standard error
+ Regression DiD with standard error(no covariates/predictors)
+ Regression DiD control for group-level predictors with standard error


+ Bayesian DiD posterior draws. posterior mean with credible interval



Does this make sense?
+ Regression DiD control for individual-level predictors

How about the Regression DiD control for group-level predictors


About $\delta_j$

+ regression DiD(TWFE regression) to get ATT and standard deviation of the noise. Then get the $\har{\delta}_j$ by subtracting individual by the group mean with standard error.

+ Bayesian DiD posterior mean. Posterior mean with credible interval.




+ number of statistical significant estimates of treatment effects
+ type S error rate. number of number of statistical significant estimates with the opposite sign
+ root mean squared error
+ rank correlation of estimates with truth


## run simulations

```{r}
# Make sure 'mod' is compiled from did_decomposed_m0plus_grouped.stan
set.seed(1)

# time trends: lambda is small
# difference between treated and control is also small
# in this case, DiD is expensive.

#cmp <- quick_compare(R = 100, N_per_state = 40, lambda = 0.5)

lambdas <- 0.3*seq_len(10)

A1s <- 0.2*seq_len(10)

cmps <- list()

for (r in seq_along(lambdas)) {
  cmps[[length(cmps) + 1L]] <- quick_compare(R = 100, N_per_state = 20,
                                             A1 = A1s[r], lambda = lambdas[r])
}


```



```{r}
print(cmp)
```





```{r}

for (r in seq_along(lambdas)){
  
  print(cmps[r]$ATT)
}
```

```{r}

lambdas <- 0.3*seq_len(10)

A1s <- 0.2*seq_len(10)

print(cmps[[1]]$ATT[1:4])
```

```{r}
print(cmps[[4]]$ATT[1:3])
print(lambdas[4])


```





```{r}
library(tibble)

# Compute true biases for ES and DM given lambda and A1
# Xbar: 2 x K matrix of state-level predictor means (row 1 = control s=0, row 2 = treated s=1)
# A1s:  either a numeric vector (when K=1), or a matrix with one row per scenario (when K>1)
# u_gap: scalar = (u_1 - u_0); default 0
true_bias_table <- function(lambdas, A1s, Xbar, u_gap = 0) {
  stopifnot(nrow(Xbar) == 2)
  deltaX <- as.numeric(Xbar[2, ] - Xbar[1, ])   # length K

  # Compute DM bias across scenarios
  if (length(deltaX) == 1L) {
    # K = 1: A1s is a numeric vector of coefficients
    dm_bias <- as.numeric(A1s) * deltaX + u_gap
    A1_label <- as.numeric(A1s)
  } else {
    # K > 1: A1s must be a matrix with rows = scenarios and K columns
    if (!is.matrix(A1s)) {
      stop("When K > 1, provide A1s as a matrix with one row per scenario and ", 
           ncol = length(deltaX), " columns.")
    }
    if (ncol(A1s) != length(deltaX)) {
      stop("ncol(A1s) must equal ncol(Xbar) when K > 1.")
    }
    dm_bias <- as.numeric(A1s %*% deltaX) + u_gap
    # Simple label: paste coefficients; you can customize as needed
    A1_label <- apply(A1s, 1, \(row) paste0("(", paste(signif(row, 3), collapse = ", "), ")"))
  }

  if (length(dm_bias) != length(lambdas)) {
    stop("Lengths of 'lambdas' and the number of A1 scenarios must match.")
  }

  tibble(
    lambda  = as.numeric(lambdas),
    A1      = A1_label,
    bias_ES = as.numeric(lambdas),  # ES bias = lambda
    bias_DM = as.numeric(dm_bias)   # DM bias = A1^T (X1 - X0) + u_gap
  )
}

# Pretty-print the first n rows (like your example)
print_true_bias <- function(lambdas, A1s, Xbar, u_gap = 0, n = 4) {
  tab <- true_bias_table(lambdas, A1s, Xbar, u_gap)
  print(head(tab, n))
  invisible(tab)
}

  
                                               
                                               

```


```{r}
lambdas <- 0.3 * seq_len(10)
A1s     <- 0.2 * seq_len(10)
Xbar    <- matrix(c(-0.5, 0.8), nrow = 2, byrow = TRUE)  # K = 1
# optional: u_gap = u1 - u0 (set to, say, 0.7 if you want DM to be biased by 0.7 even when A1=0)
u_gap   <- 0

# Compute and print first 4 rows
tab <- print_true_bias(lambdas, A1s, Xbar, u_gap, n = 10)
# If you want the full table:
# tab <- true_bias_table(lambdas, A1s, Xbar, u_gap)

```



```{r}
for (i in seq_len(10)){
  print(cmps[[i]]$ATT[1:3])
}
```

```{r}

print(cmp$DELTA_raw)
```



```{r}
# δ_j (treated counties): Bayes vs per-county DiD (aggregated)
print(cmp$DELTA)


```


