---
title: "Two_States_Two_Periods"
output: html_document
---


Two states, treated state and control state. $s =0$ for control and $s=1$ for treated. Each state has $N$ counties.
Two periods, $t=0$ not treated and $t=1$ treated.

Here is the model for $y_{jst}$. Basically, there are three key ingredients in this model: two random effects(county random effects $\gamma_j$ and time random effects $\delta$), group-level predictors $\bar{X}_{s[j]}$, and treatment effects.


The group-level predictors capture the group varying intercept and the county random effects $\gamma_j$ accounts for the individual heterogenity not explained by the group-level predictors. $\lambda$ represents the time trend we need to model. In this two-period example, $\lambda$ is just a number. In a longer panel, $\lambda_t$ is a series of numbers could be modeled by iid normal.


$$
y_{jst}= A_0+ A_1^\top \bar{X}_{s[j]} + \gamma_j + \lambda I_{t=1} + \delta_j Z_{jst} + \epsilon_{jst} 
$$

Put usual normal prior on the regression coefficients $A_1$ and intercept $A_0$. Hierarchical normal prior on $\gamma_j$ and $\gamma_j$ has different mean and variance for different groups. Normal prior on $\delta_j$. Normal prior on $\lambda$. 


$\epsilon_{jst}$ are all iid normal(0,$\sigma_{y}^2$). 


Implement five different estimates:

1. Event Study. Take difference between the mean of treated states from time $1$ and time $0$.


$\bar{y}_{s=1,t=1} - \bar{y}_{s=1, t=0} = \bar\delta + \lambda + \bar{\epsilon}_{s=1,t=1} - \bar{\epsilon}_{s=1,t=0}\sim\mathcal{N}(\bar{\delta} + \lambda, \dfrac{2\sigma_{y}^2}{N}) $

2. Difference-in-mean. Take difference between the mean of treated states from time $1$ and mean of control states from time $1$

$$
\begin{equation}
\begin{aligned}
&\bar{y}_{s=1,t=1} - \bar{y}_{s=0, t=1} 
\\ = &  \bar\delta + A_1^\top (\bar{X}_{s=1} - \bar{X}_{s=0}) + \bar{\gamma}_{s=1} - \bar{\gamma}_{s=0} + \bar{\epsilon}_{s=1,t=1} - \bar{\epsilon}_{s=0,t=1}\\
\sim & \mathcal{N}(\bar\delta + A_1^\top (\bar{X}_{s=1} - \bar{X}_{s=0}) + \bar{\gamma}_{s=1} - \bar{\gamma}_{s=0}\;,\; \dfrac{2\sigma_{y}^2}{N}) 
\end{aligned}
\end{equation}
$$


3. DiD

$$
\begin{equation}
\begin{aligned}
& (\bar{y}_{s=1,t=1} - \bar{y}_{s=1, t=0}) - (\bar{y}_{s=0,t=1} - \bar{y}_{s=0, t=0}) \\
& = \bar\delta  + \bar{\epsilon}_{s=1,t=1} - \bar{\epsilon}_{s=1,t=0} - (\bar{\epsilon}_{s=0,t=1} - \bar{\epsilon}_{s=0,t=0})\\
& \sim\mathcal{N}(\bar{\delta} , \dfrac{4\sigma_{y}^2}{N}) 
\end{aligned}
\end{equation}
$$


4. Regression DiD controlling for something need to be figured out.



5. Bayesian DiD.

We expect that conditional on the hyperparameters the Bayesian DiD estimate should be of form of the something subtracted by weighted $\lambda_t$(time trend) and $\bar{\gamma}_{s=1} - \bar{\gamma}_{s=0}$ and $ A_1^\top (\bar{X}_{s=1} - \bar{X}_{s=0})$(the group level variation).



```{r}
# =========================
# Libraries
# =========================
library(cmdstanr)
library(posterior)
library(dplyr)
library(tidyr)
library(purrr)
library(ggplot2)

# =========================
# Compile Stan once
# =========================
stan_file <- "Two_States_Two_Periods.stan"
mod <- cmdstan_model(stan_file)
```

$$
y_{jst}= A_0+ A_1^\top \bar{X}_{s[j]} + A_2^\top X_j + \gamma_j + \lambda_t + \delta_j Z_{jst} + \epsilon_{jst} 
$$


```{r simulate data}
# 2×2: one treated state (s=1), one control (s=0)
simulate_data_2x2_one_group_predictor <- function(N_per_state=40, K=1,
                                 A0=0.0, A1=0.4,
                                 Xbar = matrix(c(-0.5, 0.8), nrow=2, byrow=TRUE),
                                 u_1 = 1, u_0 = 0,
                                 sigma_c_treat=1.0, sigma_c_ctrl=0.6,
                                 # treatment & time
                                 lambda=0.5, mu_delta=3.0, sigma_delta=1.6,
                                 sigma_y=1.0, seed=1L){
  set.seed(seed)
  S <- 2L; J_t <- N_per_state; J_c <- N_per_state; J <- J_t + J_c
  is_treated <- c(rep(1L,J_t), rep(0L,J_c))
  state_of_county <- c(rep(1L,J_t), rep(2L,J_c))
  state_is_treated <- c(1L,0L)

  # truth: u_s and c_j by state group
  u <- c(u_0, u_1)
  c <- numeric(J)
  c[is_treated==1L] <- rnorm(J_t,0,sigma_c_treat)
  c[is_treated==0L] <- rnorm(J_c,0,sigma_c_ctrl)

  # deltas (treated counties only)
  delta <- rep(0,J)
  delta[is_treated==1L] <- rnorm(J_t, mu_delta, sigma_delta)

  # panel rows
  county <- rep(seq_len(J), each=2)
  post   <- rep(c(0L,1L), times=J)
  D      <- is_treated[county]*post
  s_row  <- state_of_county[county]

  A1_vec <- if (length(A1)==1) rep(A1,K) else as.numeric(A1)
  
  lin_x  <- rowSums(Xbar[s_row,,drop=FALSE] *
                    matrix(A1_vec, nrow=length(s_row), ncol=K, byrow=TRUE))
  mu <- A0 + lin_x + u[s_row] + c[county] + lambda*post + delta[county]*D
  y  <- rnorm(length(mu), mu, sigma_y)

  list(
    stan_data = list(
      N=length(y), J=J, S=2L, K=K, y=y,
      county=as.integer(county), post=as.integer(post), D=as.integer(D),
      state_of_county=as.integer(state_of_county),
      Xbar=Xbar,
      is_treated=as.integer(is_treated),
      state_is_treated=as.integer(state_is_treated),
      J_treated=J_t, treated_idx=as.integer(which(is_treated==1L)),
      center_c=0L,
      pool_u_within_group=0L,     # 2×2: no pooling across groups
      big_sd_u=10             # diffuse prior per state
    ),
    truth = list(
      A0=A0, A1=A1_vec, Xbar=Xbar, lambda=lambda, sigma_y=sigma_y,
      u=u, c=c, is_treated=is_treated, delta=delta,
      mu_delta=mu_delta, sigma_delta=sigma_delta
    )
  )
}
```


```{r}
library(cmdstanr)
library(posterior)
library(dplyr)
library(tidyr)
library(purrr)

# Classical ATT estimators (+ Wald CIs under homoskedastic σ_y)
att_classical <- function(y, county, post, state, sigma_y, N_per_state){
  y_11 <- mean(y[state==1 & post==1]); y_10 <- mean(y[state==1 & post==0])
  y_01 <- mean(y[state==0 & post==1]); y_00 <- mean(y[state==0 & post==0])
  ES  <- y_11 - y_10
  DM  <- y_11 - y_01
  DiD <- (y_11 - y_10) - (y_01 - y_00)
  z <- qnorm(0.975)
  ES_se  <- sqrt(2 * sigma_y^2 / N_per_state)
  DM_se  <- sqrt(2 * sigma_y^2 / N_per_state)
  DiD_se <- sqrt(4 * sigma_y^2 / N_per_state)
  tibble(
    method = c("ES","DM","DiD"),
    est    = c(ES, DM, DiD),
    lo     = c(ES - z*ES_se,  DM - z*DM_se,  DiD - z*DiD_se),
    hi     = c(ES + z*ES_se,  DM + z*DM_se,  DiD + z*DiD_se)
  )
}

# Per-county DiD for treated counties (analytic CI under σ_y)
per_county_did <- function(y, county, post, state, sigma_y, N_per_state, treated_ids){
  y_c1 <- mean(y[state==0 & post==1]); y_c0 <- mean(y[state==0 & post==0])
  y_first_difference_control <- y_c1 - y_c0
  z <- qnorm(0.975)
  se_dj <- sqrt(2*sigma_y^2 + 2*sigma_y^2/N_per_state)
  map_dfr(treated_ids, function(j){
    y_j1 <- y[county==j & post==1]; y_j0 <- y[county==j & post==0]
    dj <- (y_j1 - y_j0) - y_first_difference_control
    tibble(county=j, est=dj, lo=dj - z*se_dj, hi=dj + z*se_dj)
  })
}
```

# Bayesian DiD via cmdstanr

```{r}
# ---- one-time: pick a durable folder for chain CSVs ----
.default_outdir <- file.path(getwd(), "cmdstan_out")
if (!dir.exists(.default_outdir)) dir.create(.default_outdir, recursive = TRUE)

fit_bayes_once <- function(stan_data,
                           treated_idx  = NULL, # will try stan_data$treated_idx if NULL
                           delta_name   = "delta",
                           iter_warmup  = 500,
                           iter_sampling= 1000,
                           chains       = 4,
                           parallel_chains = chains,
                           seed         = 2025,
                           output_dir   = .default_outdir,
                           refresh      = 0) {
  
  # --- resolve treated indices ---
  if (is.null(treated_idx)) {
    treated_idx <- stan_data$treated_idx %||% stan_data$treated_ids
    if (is.null(treated_idx)) stop("Provide treated_idx or include $treated_idx in stan_data.")
  }
  treated_idx <- as.integer(treated_idx)
  J1 <- length(treated_idx)

  # --- unique chain ids to avoid filename collisions ---
  chain_ids <- seq_len(chains)

  # --- run sampler with pinned output_dir ---
  
  outdir <- file.path(getwd(), "cmdstan_out")
  dir.create(outdir, showWarnings = FALSE, recursive = TRUE)
  
  stopifnot(all(is.finite(unlist(stan_data))))

  fit <- mod$sample(
    data = stan_data,
    chains = chains, parallel_chains = chains,
    iter_warmup = iter_warmup, iter_sampling = iter_sampling,
    refresh = 0, show_messages = FALSE,output_dir = outdir,
  save_cmdstan_config = TRUE,   # <- key
  save_warmup = TRUE,           # optional
  output_basename = "run1",    # optional: stable filenames
  seed = seed
  )

  # --- check CSVs exist & chains returned 0 ---
  csvs <- fit$output_files(include_failed = TRUE)
  if (!all(file.exists(csvs))) {
    message("Some chain CSVs are missing:\n", paste(csvs[!file.exists(csvs)], collapse = "\n"))
  }
  rc <- fit$return_codes()
  if (any(rc != 0L) || !all(file.exists(csvs))) {
    # print stderr/stdout paths to help debug
    std_out <- fit$stdout_files()
    diag_files <- fit$diagnostic_files()
    message("Nonzero return codes: ", paste(rc, collapse = ", "))
    message("Stdout files:\n", paste(std_out, collapse = "\n"))
    message("Diagnostic files:\n", paste(diag_files, collapse = "\n"))
    # Try a safer rerun (single chain) to surface the underlying error
    message("Retrying with chains=1, parallel_chains=1, refresh=0 to surface messages...")
    
    stopifnot(all(is.finite(unlist(stan_data))))

    fit <- mod$sample(
      data = stan_data,
      seed = seed + 1L,
      chains = 1,
      parallel_chains = 1,
      chain_ids = 1,
      iter_warmup = iter_warmup,
      iter_sampling = iter_sampling,
      refresh = 0,
      show_messages = FALSE,
      output_dir = output_dir
    )
  }

  # --- extract delta draws ONLY (avoid delta_raw) ---
  adapt_delta = 0.9
  
  dr_delta <- posterior::as_draws_df(fit$draws(variables = delta_name))
  delta_cols <- grep(paste0("^", delta_name, "\\[[0-9]+\\]$"),
                     names(dr_delta), value = TRUE)
  if (length(delta_cols) == 0) {
    stop("No columns found for parameter '", delta_name, "'. Check Stan variable names.")
  }

  j_stan <- as.integer(sub(paste0("^", delta_name, "\\[([0-9]+)\\]$"), "\\1", delta_cols))
  ord <- order(j_stan)
  delta_cols <- delta_cols[ord]; j_stan <- j_stan[ord]
  Delta_mat <- as.matrix(dr_delta[, delta_cols, drop = FALSE])

  # --- handle two cases: delta over ALL counties vs TREATED-only ---
  if (!is.null(stan_data$J) && length(j_stan) == stan_data$J) {
    pos_treat <- match(treated_idx, j_stan)
    if (anyNA(pos_treat)) {
      stop("Some treated_idx are not present among indices of '", delta_name, "'.")
    }
    Delta_tr <- Delta_mat[, pos_treat, drop = FALSE]
    county_labels <- treated_idx
  } else if (length(j_stan) == J1 && all(j_stan == seq_len(J1))) {
    # Stan stored delta only for treated counties
    Delta_tr <- Delta_mat
    county_labels <- treated_idx
  } else {
    stop("Dimension mismatch: ", delta_name, " has ", length(j_stan),
         " entries; expected J (all) or J1 (treated).")
  }

  # --- per-county posteriors (treated only) ---
  delta_mean <- colMeans(Delta_tr)
  delta_sd   <- apply(Delta_tr, 2, stats::sd)
  delta_l95  <- apply(Delta_tr, 2, stats::quantile, probs = 0.025)
  delta_u95  <- apply(Delta_tr, 2, stats::quantile, probs = 0.975)
  

  delta_tbl <- tibble::tibble(
    j   = as.integer(county_labels),
    est = as.numeric(delta_mean),
    sd  = as.numeric(delta_sd),
    lo  = as.numeric(delta_l95),
    hi  = as.numeric(delta_u95)
  )

  # --- ATT from the same draws (coherent) ---
  ATT_draws <- rowMeans(Delta_tr)
  ATT_mean  <- mean(ATT_draws)
  ATT_ci    <- stats::quantile(ATT_draws, c(0.025, 0.975), names = FALSE)

  
  list(
    ATT = list(mean = ATT_mean, lo = ATT_ci[1], hi = ATT_ci[2], draws = ATT_draws),
    delta = delta_tbl,
    output_files = fit$output_files()
  )
}
```


```{r}
quick_compare <- function(R = 50, N_per_state = 40, K = 1,
                          A0 = 0.0, A1 = 0.4,
                          Xbar = matrix(c(-0.5, 0.8), nrow = 2, byrow = TRUE),
                          u_1 = 1, u_0 = 0,
                          sigma_c_treat = 1.0, sigma_c_ctrl = 0.7,
                          lambda = 2, mu_delta = 1.0, sigma_delta = 0.6,
                          sigma_y = 1.0, seed0 = 1234,
                          iter_warmup = 500, iter_sampling = 500, chains = 2) {
  # Storage for results. Lists of data.frames/tibbles
  ATT_rows   <- vector("list", R)
  DELTA_rows <- vector("list", R)
  for (r in seq_len(R)) {
    sim <- simulate_data_2x2_one_group_predictor(N_per_state, K, A0, A1, Xbar,
                                u_1, u_0, sigma_c_treat, sigma_c_ctrl,
                                lambda, mu_delta, sigma_delta, sigma_y,
                              seed = seed0 + r)
    ## classical (ATT + per-county DiD)
    county <- rep(seq_len(sim$stan_data$J), each = 2)
    post   <- sim$stan_data$post
    
    # sim$stan_data$state_of_county is 1-based; convert to 0-based for R indexing
    state  <- sim$stan_data$state_of_county - 1L
    y      <- sim$stan_data$y
    Nps    <- N_per_state
    treated_ids <- which(sim$truth$is_treated == 1L)

    
    
    # to use functions att_classical and per_county_did, we need to estimate sigma_y. Just to remark, att_classical and per_county_did assume homoskedasticity with estimated sigma_y. Without homoskedasticity assumption, there is no way to estimate the variance of treatment effect estimators for each county. The analytic formula is specific to this particular model.
    
    twfe_model <- lm(y ~ factor(county) + factor(post) + factor(state):factor(post))
    
    sigma_did_estimate <- summary(twfe_model)$sigma
    
    sigma_y_estimated <- sigma_did_estimate/2*sqrt(N_per_state)
    
    cls  <- att_classical(y, county, post, state, sigma_y_estimated, Nps)   
    
    # data.frame with {method, est, lo, hi}
    dj   <- per_county_did(y, county, post, state, sigma_y_estimated, Nps, treated_ids)      # tibble {county, est, lo, hi}

    ## Bayes (per-county posteriors + ATT from the same draws)
    fitb <- fit_bayes_once(
      stan_data       = sim$stan_data,
      treated_idx     = sim$stan_data$treated_idx,
      iter_warmup     = iter_warmup,
      iter_sampling   = iter_sampling,
      chains          = chains,
      parallel_chains = chains,
      seed            = seed0 + 10000 + r,
      refresh         = 0
    )

    ## --- ATT metrics ---
    true_delta <- sim$truth$delta
    true_ATT <- mean(sim$truth$delta[treated_ids])
    
    att_all <- dplyr::bind_rows(
      cls,
      tibble::tibble(method = "Bayes DiD", est = fitb$ATT$mean, lo = fitb$ATT$lo, hi = fitb$ATT$hi)
    ) |>
      dplyr::mutate(true = true_ATT,
                    bias = est - true,
                    squared_error = (est - true)^2
                    #overed = as.integer(true >= lo & true <= hi),
                    #width = hi - lo)
      )
    ATT_rows[[r]] <- att_all

    ## --- δ_j metrics (treated counties): DiD vs Bayes ---
    
    delta_bayes_did <- fitb$delta |>
      dplyr::filter(j %in% treated_ids) |>
      dplyr::rename(est_bayes_did = est, lo_bayes_did = lo, hi_bayes_did = hi)
    
    delta_did <- dj |>
      dplyr::rename(j = county, est_did = est, lo_did = lo, hi_did = hi)
    
    # inner join by j
    
    d_join <- dplyr::inner_join(delta_bayes_did, delta_did, by = "j") |>
      dplyr::mutate(
        truth = sim$truth$delta[j],
        
        ## significance (two-sided 5%).
        
        sig_bayes_did = ( 0 > hi_bayes_did ) | (0 < lo_bayes_did),
        sig_did = (0 > hi_did  ) | (0 < lo_did),
        
        ## Type-S indicators among the significant
        
        typeS_bayes_did = sig_bayes_did & (sign(est_bayes_did) != sign(truth)),
        typeS_did = sig_did & (sign(est_did) != sign(truth)),
        
        ## squared errors
        squared_error_bayes_did = (est_bayes_did - truth)^2,
        squared_error_did = (est_did - truth)^2
      )
    n_sig_bayes_did <- sum(d_join$sig_bayes_did)
    n_sig_did <- sum(d_join$sig_did)
    typeS_cnt_b <- sum(d_join$typeS_bayes_did)
    typeS_cnt_f <- sum(d_join$typeS_did)

    DELTA_rows[[r]] <- tibble::tibble(
      ## Counts
      
      n_sig_bayes_did = n_sig_bayes_did,
      n_sig_did = n_sig_did,
      ## Type-S counts and rates (NA if no significant)
      typeS_bayes_did_count = typeS_cnt_b,
      typeS_did_count = typeS_cnt_f,
      typeS_bayes_did_rate  = ifelse(n_sig_bayes_did > 0, typeS_cnt_b / n_sig_bayes_did, NA_real_),
      typeS_did_rate  = ifelse(n_sig_did > 0, typeS_cnt_f / n_sig_did, NA_real_),
      ## MSE and RMSE
      mse_b   = mean(d_join$squared_error_bayes_did),
      mse_f   = mean(d_join$squared_error_did),
      rmse_b  = sqrt(mean(d_join$squared_error_bayes_did)),
      rmse_f  = sqrt(mean(d_join$squared_error_did)),
      ## Coverage and average width (kept for completeness)
      #cover_b = mean(d_join$truth >= db$lo_bayes_did & d_join$truth <= db$hi_bayes_did),
      #cover_f = mean(d_join$truth >= df$lo_did & d_join$truth <= df$hi_did),
      #width_b = mean(db$hi_bayes_did - db$lo_bayes_did),
      #width_f = mean(df$hi_did - df$lo_did),
      ## Spearman rank correlations
      rank_b  = suppressWarnings(cor(d_join$est_bayes_did, d_join$truth, method = "spearman")),
      rank_f  = suppressWarnings(cor(d_join$est_did, d_join$truth, method = "spearman"))
    )
  }

  ATT_all   <- dplyr::bind_rows(ATT_rows)
  DELTA_all <- dplyr::bind_rows(DELTA_rows)

  att_sum <- ATT_all |>
    dplyr::group_by(method) |>
    dplyr::summarize(
      rmse = sqrt(mean(squared_error)),
      mse  = mean(squared_error),
      bias = mean(bias),
      #cover95 = mean(covered),
      #avg_width = mean(width),
      .groups = "drop"
    )

  delta_sum <- DELTA_all |>
    dplyr::summarize(
      ## Bayes
      n_sig_bayes_did_mean = mean(n_sig_bayes_did),
      typeS_bayes_did_count_mean = mean(typeS_bayes_did_count),
      typeS_bayes_did_rate_mean  = mean(typeS_bayes_did_rate, na.rm = TRUE),
      mse_b = mean(mse_b), rmse_b = mean(rmse_b),
      #cover_b = mean(cover_b), width_b = mean(width_b),
      rank_b  = mean(rank_b, na.rm = TRUE),
      ## DiD
      n_sig_did_mean = mean(n_sig_did),
      typeS_did_count_mean = mean(typeS_did_count),
      typeS_did_rate_mean  = mean(typeS_did_rate, na.rm = TRUE),
      mse_f = mean(mse_f), rmse_f = mean(rmse_f),
      #cover_f = mean(cover_f), width_f = mean(width_f),
      rank_f  = mean(rank_f, na.rm = TRUE)
    )

  list(ATT = att_sum,
       DELTA = delta_sum,
       ATT_raw = ATT_all,
       DELTA_raw = DELTA_all)
}
```



```{r}
res$DELTA_raw
```




```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(rlang)

# ---------- Column harmonizer ----------
harmonize_delta_cols <- function(df) {
  first_hit <- function(df, candidates, new) {
    hit <- intersect(candidates, names(df))
    if (length(hit) && !(new %in% names(df))) df[[new]] <<- df[[hit[1]]]
  }
  # MSE / RMSE
  first_hit(df, c("mse_b","mse_BDiD","mse_bayes_did","MSE_b","MSE_BDiD"), "mse_b")
  first_hit(df, c("mse_f","mse_DiD","mse_freq","MSE_f","MSE_DiD"),       "mse_f")
  first_hit(df, c("rmse_b","RMSE_b","rmse_BDiD"),                         "rmse_b")
  first_hit(df, c("rmse_f","RMSE_f","rmse_DiD"),                          "rmse_f")
  # Rank
  first_hit(df, c("rank_b","spearman_b","rho_b"),                         "rank_b")
  first_hit(df, c("rank_f","spearman_f","rho_f"),                         "rank_f")
  # Sig counts
  first_hit(df, c("n_sig_bayes_did_mean","n_sig_b_mean","n_sig_BDiD"),    "n_sig_bayes_did_mean")
  first_hit(df, c("n_sig_did_mean","n_sig_f_mean","n_sig_DiD"),           "n_sig_did_mean")
  # Type-S
  first_hit(df, c("typeS_bayes_did_count_mean","typeS_b_count_mean"),     "typeS_bayes_did_count_mean")
  first_hit(df, c("typeS_did_count_mean","typeS_f_count_mean"),           "typeS_did_count_mean")
  first_hit(df, c("typeS_bayes_did_rate_mean","typeS_b_rate_mean"),       "typeS_bayes_did_rate_mean")
  first_hit(df, c("typeS_did_rate_mean","typeS_f_rate_mean"),             "typeS_did_rate_mean")
  df
}

# Helper: pick x var safely ("lambda" or "u_1")
.pick_x <- function(df, x_axis) {
  stopifnot(x_axis %in% c("lambda","u_1"))
  if (!(x_axis %in% names(df))) {
    abort(paste0("Column '", x_axis, "' not found in data."))
  }
  x_axis
}

# Helper: add facet if requested (facet_by is a STRING or NULL)
.add_facet <- function(p, facet_by) {
  if (!is.null(facet_by)) {
    if (!facet_by %in% colnames(p$data)) {
      abort(paste0("Facet column '", facet_by, "' not found in data."))
    }
    p + facet_wrap(vars(!!sym(facet_by)), scales = "free_y")
  } else p
}

# ---------- 1) MSE ----------
plot_delta_mean_squared_error <- function(grid_out, x_axis = "lambda", facet_by = NULL){
  df <- harmonize_delta_cols(grid_out)
  x_axis <- .pick_x(df, x_axis)
  need <- c("mse_b","mse_f", x_axis)
  if (!all(need %in% names(df))) abort("Missing columns: ", paste(setdiff(need, names(df)), collapse=", "))
  plot_df <- df |>
    select(all_of(c(x_axis, facet_by)), mse_b, mse_f) |>
    pivot_longer(cols = c(mse_b, mse_f), names_to = "method", values_to = "value") |>
    mutate(method = ifelse(method == "mse_b", "Bayes", "DiD"))
  p <- ggplot(plot_df, aes(x = .data[[x_axis]], y = value, linetype = method, shape = method)) +
    geom_line(linewidth = 0.9) + geom_point(size = 1.8) +
    scale_linetype_manual(values = c(Bayes = "dashed", DiD = "solid")) +
    scale_shape_manual(values = c(Bayes = 17, DiD = 16)) +
    labs(x = if (x_axis == "lambda") expression(lambda) else expression(u[1]),
         y = "MSE", linetype = "Method", shape = "Method") +
    theme_minimal(base_size = 12) + theme(legend.position = "bottom")
  .add_facet(p, facet_by)
}

# ---------- 2) Rank correlation ----------
plot_delta_rank_correlation <- function(grid_out, x_axis = "lambda", facet_by = NULL) {
  df <- harmonize_delta_cols(grid_out)
  x_axis <- .pick_x(df, x_axis)
  need <- c("rank_b","rank_f", x_axis)
  if (!all(need %in% names(df))) abort("Missing columns: ", paste(setdiff(need, names(df)), collapse=", "))
  plot_df <- df |>
    select(all_of(c(x_axis, facet_by)), rank_b, rank_f) |>
    pivot_longer(cols = c(rank_b, rank_f), names_to = "method", values_to = "value") |>
    mutate(method = ifelse(method == "rank_b", "Bayes", "DiD"))
  p <- ggplot(plot_df, aes(x = .data[[x_axis]], y = value, linetype = method, shape = method)) +
    geom_line(linewidth = 0.9) + geom_point(size = 1.8) +
    scale_linetype_manual(values = c(Bayes = "dashed", DiD = "solid")) +
    scale_shape_manual(values = c(Bayes = 17, DiD = 16)) +
    labs(x = if (x_axis == "lambda") expression(lambda) else expression(u[1]),
         y = "Spearman ρ", linetype = "Method", shape = "Method") +
    theme_minimal(base_size = 12) + theme(legend.position = "bottom")
  .add_facet(p, facet_by)
}

# ---------- 3) Number significant ----------
plot_delta_significant_counts <- function(grid_out, x_axis = "lambda", facet_by = NULL) {
  df <- harmonize_delta_cols(grid_out)
  x_axis <- .pick_x(df, x_axis)
  need <- c("n_sig_bayes_did_mean","n_sig_did_mean", x_axis)
  if (!all(need %in% names(df))) abort("Missing columns: ", paste(setdiff(need, names(df)), collapse=", "))
  plot_df <- df |>
    select(all_of(c(x_axis, facet_by)), n_sig_bayes_did_mean, n_sig_did_mean) |>
    pivot_longer(cols = c(n_sig_bayes_did_mean, n_sig_did_mean),
                 names_to = "method", values_to = "value") |>
    mutate(method = ifelse(method == "n_sig_bayes_did_mean", "Bayes", "DiD"))
  p <- ggplot(plot_df, aes(x = .data[[x_axis]], y = value, linetype = method, shape = method)) +
    geom_line(linewidth = 0.9) + geom_point(size = 1.8) +
    scale_linetype_manual(values = c(Bayes = "dashed", DiD = "solid")) +
    scale_shape_manual(values = c(Bayes = 17, DiD = 16)) +
    labs(x = if (x_axis == "lambda") expression(lambda) else expression(u[1]),
         y = "Number significant (5%)", linetype = "Method", shape = "Method") +
    theme_minimal(base_size = 12) + theme(legend.position = "bottom")
  .add_facet(p, facet_by)
}

# ---------- 4) Type-S (count or rate) ----------
plot_delta_type_S_errors <- function(grid_out, x_axis = "lambda", facet_by = NULL,
                                     use_typeS = c("count","rate")) {
  use_typeS <- match.arg(use_typeS)
  df <- harmonize_delta_cols(grid_out)
  x_axis <- .pick_x(df, x_axis)

  if (use_typeS == "count") {
    need <- c("typeS_bayes_did_count_mean","typeS_did_count_mean", x_axis)
    if (!all(need %in% names(df))) abort("Missing columns: ", paste(setdiff(need, names(df)), collapse=", "))
    plot_df <- df |>
      select(all_of(c(x_axis, facet_by)), typeS_bayes_did_count_mean, typeS_did_count_mean) |>
      pivot_longer(cols = c(typeS_bayes_did_count_mean, typeS_did_count_mean),
                   names_to = "method", values_to = "value") |>
      mutate(method = ifelse(method == "typeS_bayes_did_count_mean", "Bayes", "DiD"))
    ylab <- "Mean Type-S (count)"
  } else {
    need <- c("typeS_bayes_did_rate_mean","typeS_did_rate_mean", x_axis)
    if (!all(need %in% names(df))) abort("Missing columns: ", paste(setdiff(need, names(df)), collapse=", "))
    plot_df <- df |>
      select(all_of(c(x_axis, facet_by)), typeS_bayes_did_rate_mean, typeS_did_rate_mean) |>
      pivot_longer(cols = c(typeS_bayes_did_rate_mean, typeS_did_rate_mean),
                   names_to = "method", values_to = "value") |>
      mutate(method = ifelse(method == "typeS_bayes_did_rate_mean", "Bayes", "DiD"))
    ylab <- "Type-S rate"
  }

  p <- ggplot(plot_df, aes(x = .data[[x_axis]], y = value, linetype = method, shape = method)) +
    geom_line(linewidth = 0.9) + geom_point(size = 1.8) +
    scale_linetype_manual(values = c(Bayes = "dashed", DiD = "solid")) +
    scale_shape_manual(values = c(Bayes = 17, DiD = 16)) +
    labs(x = if (x_axis == "lambda") expression(lambda) else expression(u[1]),
         y = ylab, linetype = "Method", shape = "Method") +
    theme_minimal(base_size = 12) + theme(legend.position = "bottom")
  .add_facet(p, facet_by)
}
```






```{r}
# draw plots

p1 <- plot_delta_mean_squared_error(grid_out)

p2 <- plot_delta_rank_correlation(grid_out)

p3 <- plot_delta_significant_counts(grid_out)

p4 <- plot_delta_type_S_errors(grid_out, use_typeS = "rate")

library(patchwork)
(p1 | p2) / (p3 | p4) + plot_layout(guides = "collect") & theme(legend.position = 'bottom')


```




```{r}
library(purrr); library(dplyr); library(tidyr)

run_grid <- function(lambdas, u1_grid, R = 50, N_per_state = 40, ...) {
  map2_dfr(lambdas, u1_grid, \(lam, u1) {
    res <- quick_compare(R = R, N_per_state = N_per_state, lambda = lam, u_1 = u1,  ...)
    # Flatten: keep only delta summaries for plotting the 4 quantities
    tibble(lambda = lam, u_1 = u1) |>
      bind_cols(res$DELTA)
  })
}

```




```{r}
# Example:
# record the run time
# run R = 1, lambdas < -seq(0,1,by=1), two lambdas, iter_warmup 300, iter_sampling = 600, chains = 1
# takes 16 secs


# run R = 1, lambdas < -seq(0, 9, by=1), 10 lambdas, iter_warmup 300, iter_sampling = 600, chains = 1
# takes 83.9 secs


# run R = 50, lambdas < -seq(0, 9, by=1), 10 lambdas, iter_warmup 300, iter_sampling = 600, chains = 1
# takes 83.9 secs*50 secs = 4195 secs = 70 mins
# diagnostics. One reports ER < 0.3. We need to run more iterations int the warmup.


# run R = 20, lambdas <- seq(0, 20, by = 3), 7 lambdas, iter_warmup 500, iter_sampling = 600, chains=1
# 25 mins
# run R = 10, lambdas <- seq(0, 21, by = 7), 4 lambdas, iter_warmup 500, iter_sampling = 600, chains=2
# 10 mins

start_time <- Sys.time()
lambdas <- seq(0, 21, by = 7)
u1_grid <-  rep(1, length(lambdas))  
grid_out_lambda_varying <- run_grid(lambdas, u1_grid, R = 10, iter_warmup = 500, iter_sampling = 600, chains = 2)

end_time <- Sys.time()
end_time - start_time

```



```{r}
# print some results of grid_out

print(grid_out_lambda_varying)

# save grid_out_lambda_varying to local file

saveRDS(grid_out_lambda_varying, file = "data/grid_out_lambda_varying.rds")

p1_lambda_varying <- plot_delta_mean_squared_error(grid_out_lambda_varying)

p2_lambda_varying <- plot_delta_rank_correlation(grid_out_lambda_varying)

p3_lambda_varying <- plot_delta_significant_counts(grid_out_lambda_varying)

p4_lambda_varying <- plot_delta_type_S_errors(grid_out_lambda_varying, use_typeS = "rate")

(p1_lambda_varying | p2_lambda_varying) / (p3_lambda_varying | p4_lambda_varying) + plot_layout(guides = "collect") & theme(legend.position = 'bottom')

# save the plots to local file

ggsave("images/grid_out_lambda_varying_plots.png", width = 10, height = 8)

```



```{r}

# Example: hold lambda and u0 fixed and vary the difference between u1 and u0. Just vary u1. Set u0 = 0.

u1_grid  <- seq(1, 25, by = 7)  
lambdas <- rep(2, length(u1_grid))
grid_out_u1_varying <- run_grid(lambdas, u1_grid, R = 10, iter_warmup = 500, iter_sampling = 600, chains = 2)
```


```{r}
# print some results of grid_out
print(grid_out_u1_varying)

saveRDS(grid_out_u1_varying, file = "data/grid_out_u1_varying.rds")

# plots for delta_j estimates

p1_u1_varying <- plot_delta_mean_squared_error(grid_out_u1_varying, x_axis = "u_1")

p2_u1_varying <- plot_delta_rank_correlation(grid_out_u1_varying, x_axis = "u_1")

p3_u1_varying <- plot_delta_significant_counts(grid_out_u1_varying, x_axis = "u_1")

p4_u1_varying <- plot_delta_type_S_errors(grid_out_u1_varying, x_axis = "u_1", use_typeS = "rate")

(p1_u1_varying | p2_u1_varying) / (p3_u1_varying | p4_u1_varying) + plot_layout(guides = "collect") & theme(legend.position = 'bottom')

ggsave("images/grid_out_u1_varying_plots.png", width = 10, height = 8)

```


```{r more grids few replicates simulations}
start_time <- Sys.time()
lambdas_large <- seq(-100, 100, by = 10)
u1_grid_large <-  rep(1, length(lambdas_large))  
grid_out_lambda_varying_large <- run_grid(lambdas_large, u1_grid_large, R = 10, iter_warmup = 800, iter_sampling = 600, chains = 2)
end_time <- Sys.time()
end_time - start_time
```

### decrease the sigma_y to make it more favorable to DiD and see how our approach performs


```{r}

start_time <- Sys.time()
lambdas_large <- seq(-100, 100, by = 25)
u1_grid_large <-  rep(1, length(lambdas_large))  
grid_out_lambda_varying_large_small_sigma_y <- run_grid(lambdas_large, u1_grid_large, sigma_y = 0.2, R = 10, iter_warmup = 800, iter_sampling = 600, chains = 2)
end_time <- Sys.time()
end_time - start_time

```


```{r}
saveRDS(grid_out_lambda_varying_large_small_sigma_y, file = "data/grid_out_lambda_varying_large_small_sigma_y.rds")

p1_lambda_varying_large_small_sigma_y <- plot_delta_mean_squared_error(grid_out_lambda_varying_large_small_sigma_y, x_axis = "lambda")

p2_lambda_varying_large_small_sigma_y <- plot_delta_rank_correlation(grid_out_lambda_varying_large_small_sigma_y, x_axis = "lambda")

p3_lambda_varying_large_small_sigma_y <- plot_delta_significant_counts(grid_out_lambda_varying_large_small_sigma_y, x_axis = "lambda")

p4_lambda_varying_large_small_sigma_y <- plot_delta_type_S_errors(grid_out_lambda_varying_large_small_sigma_y,x_axis = "lambda" , use_typeS = "rate")

(p1_lambda_varying_large_small_sigma_y | p2_lambda_varying_large_small_sigma_y) / (p3_lambda_varying_large_small_sigma_y | p4_lambda_varying_large_small_sigma_y) + plot_layout(guides = "collect") & theme(legend.position = 'bottom')

ggsave("images/grid_out_lambda_varying_large_plots_small_sigma_y.png", width = 10, height = 8)

```

```{r}

u1_grid  <- seq(-100, 100, by = 10)  
lambdas <- rep(2, length(u1_grid))
grid_out_u1_varying_large_small_sigma_y <- run_grid(lambdas, u1_grid,sigma_y = 0.2, R = 30, iter_warmup = 800, iter_sampling = 600, chains = 2)

# print some results of grid_out_u1_varying_large
print(grid_out_u1_varying_large_small_sigma_y)

saveRDS(grid_out_u1_varying_large_small_sigma_y, file = "data/grid_out_u1_varying_large_small_sigma_y.rds")

p1_u1_varying_large_small_sigma_y <- plot_delta_mean_squared_error(grid_out_u1_varying_large_small_sigma_y, x_axis = "u_1")

p2_u1_varying_large_small_sigma_y <- plot_delta_rank_correlation(grid_out_u1_varying_large_small_sigma_y, x_axis = "u_1")

p3_u1_varying_large_small_sigma_y <- plot_delta_significant_counts(grid_out_u1_varying_large_small_sigma_y, x_axis = "u_1")

p4_u1_varying_large_small_sigma_y <- plot_delta_type_S_errors(grid_out_u1_varying_large_small_sigma_y, use_typeS = "rate", x_axis = "u_1")

(p1_u1_varying_large_small_sigma_y | p2_u1_varying_large_small_sigma_y) / (p3_u1_varying_large_small_sigma_y | p4_u1_varying_large_small_sigma_y) + plot_layout(guides = "collect") & theme(legend.position = 'bottom')

ggsave("images/grid_out_u1_varying_large_plots_small_sigma_y.png", width = 10, height = 8)

```









### Run the u for the same range -100 to 100

### Try to decrease the sigma_y to make it more favorable to DiD and see how our approach performs




```{r}
print(grid_out_lambda_varying_large)

saveRDS(grid_out_lambda_varying_large, file = "data/grid_out_lambda_varying_large_wider_range.rds")

p1_lambda_varying_large <- plot_delta_mean_squared_error(grid_out_lambda_varying_large, x_axis = "lambda")

p2_lambda_varying_large <- plot_delta_rank_correlation(grid_out_lambda_varying_large, x_axis = "lambda")

p3_lambda_varying_large <- plot_delta_significant_counts(grid_out_lambda_varying_large, x_axis = "lambda")

p4_lambda_varying_large <- plot_delta_type_S_errors(grid_out_lambda_varying_large,x_axis = "lambda" , use_typeS = "rate")
```


```{r}
library(patchwork)
(p1_lambda_varying_large | p2_lambda_varying_large) / (p3_lambda_varying_large | p4_lambda_varying_large) + plot_layout(guides = "collect") & theme(legend.position = 'bottom')

ggsave("images/grid_out_lambda_varying_large_plots_wider_range.png", width = 10, height = 8)
```


```{r}

mu_delta<-3
sigma_delta<-1.6

delta <- rnorm(40, mu_delta, sigma_delta)

# count the statistical significant delta estimates suppose sigma_delta is known

sum((delta - 0)/sigma_delta > 1.96 | (delta - 0)/sigma_delta < -1.96)

# replicate many times

set.seed(1234)
R <- 1000
sig_counts <- replicate(R, {
  delta <- rnorm(40, mu_delta, sigma_delta)
  sum((delta - 0)/sigma_delta > 1.96 | (delta - 0)/sigma_delta < -1.96)
})

# mark the mean significant counts

mean(sig_counts)
hist(sig_counts, breaks = seq(-0.5, 40.5, by = 1), main = "Histogram of Significant Counts", xlab = "Number of Significant Counts")
abline(v = mean(sig_counts), col = "red", lwd = 2)
#show the mean at x-axis

# move the text a bit to the top left

text(x = mean(sig_counts) -8, y = 120, labels = paste("Mean =", round(mean(sig_counts),2)), col = "red")

# show median

median_sig <- median(sig_counts)
abline(v = median_sig, col = "blue", lwd = 2)
text(x = median_sig + 7, y = 120, labels = paste("Median =", round(median_sig,2)), col = "blue")



saveRDS(sig_counts, file = "data/sig_counts_simulation.rds")

```


```{r}

# Open a PNG device
png("images/sig_counts_histogram.png", width = 800, height = 600)

hist(sig_counts, 
     breaks = seq(-0.5, 40.5, by = 1), 
     main = "Histogram of Significant Counts", 
     xlab = "Number of Significant Counts")

# Add mean line and label
abline(v = mean(sig_counts), col = "red", lwd = 2)
text(x = mean(sig_counts) - 8, y = 120, 
     labels = paste("Mean =", round(mean(sig_counts), 2)), 
     col = "red")

# Add median line and label
abline(v = median_sig, col = "blue", lwd = 2)
text(x = median_sig + 7, y = 120, 
     labels = paste("Median =", round(median_sig, 2)), 
     col = "blue")

# Close the PNG device
dev.off()

# Save as PDF as well
pdf("images/sig_counts_histogram.pdf", width = 8, height = 6)

hist(sig_counts, 
     breaks = seq(-0.5, 40.5, by = 1), 
     main = "Histogram of Significant Counts", 
     xlab = "Number of Significant Counts")

abline(v = mean(sig_counts), col = "red", lwd = 2)
text(x = mean(sig_counts) - 8, y = 120, 
     labels = paste("Mean =", round(mean(sig_counts), 2)), 
     col = "red")

abline(v = median_sig, col = "blue", lwd = 2)
text(x = median_sig + 7, y = 120, 
     labels = paste("Median =", round(median_sig, 2)), 
     col = "blue")

dev.off()


```




```{r}
u1_grid  <- seq(-100, 100, by = 10)  
lambdas <- rep(2, length(u1_grid))
grid_out_u1_varying_large <- run_grid(lambdas, u1_grid, R = 10, iter_warmup = 800, iter_sampling = 600, chains = 2)
```

```{r}
# print some results of grid_out_u1_varying_large
print(grid_out_u1_varying_large)

saveRDS(grid_out_u1_varying_large, file = "data/grid_out_u1_varying_large.rds")

p1_u1_varying_large <- plot_delta_mean_squared_error(grid_out_u1_varying_large, x_axis = "u_1")

p2_u1_varying_large <- plot_delta_rank_correlation(grid_out_u1_varying_large, x_axis = "u_1")

p3_u1_varying_large <- plot_delta_significant_counts(grid_out_u1_varying_large, x_axis = "u_1")

p4_u1_varying_large <- plot_delta_type_S_errors(grid_out_u1_varying_large, use_typeS = "rate", x_axis = "u_1")

(p1_u1_varying_large | p2_u1_varying_large) / (p3_u1_varying_large | p4_u1_varying_large) + plot_layout(guides = "collect") & theme(legend.position = 'bottom')

ggsave("images/grid_out_u1_varying_large_plots.png", width = 10, height = 8)
```




```{r more grids few replicates simulations}

start_time <- Sys.time()
lambdas_large <- seq(0, 21, by = 3)
u1_grid_large <-  rep(1, length(lambdas))  
grid_out_lambda_varying_large <- run_grid(lambdas, u1_grid, R = 50, iter_warmup = 1000, iter_sampling = 600, chains = 2)

end_time <- Sys.time()
end_time - start_time


print(grid_out_lambda_varying)

p1_lambda_varying_large <- plot_delta_mean_squared_error(grid_out_lambda_varying_large)

p2_lambda_varying_large <- plot_delta_rank_correlation(grid_out_lambda_varying_large)

p3_lambda_varying_large <- plot_delta_significant_counts(grid_out_lambda_varying_large)

p4_lambda_varying_large <- plot_delta_type_S_errors(grid_out_lambda_varying_large, use_typeS = "rate")

(p1_lambda_varying_large | p2_lambda_varying_large) / (p3_lambda_varying_large | p4_lambda_varying_large) + plot_layout(guides = "collect") & theme(legend.position = 'bottom')


u1_grid  <- seq(1, 25, by = 3)  
lambdas <- rep(2, length(u1_grid))
grid_out_u1_varying_large <- run_grid(lambdas, u1_grid, R = 50, iter_warmup = 1000, iter_sampling = 600, chains = 2)

# print some results of grid_out
print(grid_out_u1_varying_large)

p1_u1_varying_large <- plot_delta_mean_squared_error(grid_out_u1_varying_large)

p2_u1_varying_large <- plot_delta_rank_correlation(grid_out_u1_varying_large)

p3_u1_varying_large <- plot_delta_significant_counts(grid_out_u1_varying_large)

p4_u1_varying_large <- plot_delta_type_S_errors(grid_out_u1_varying_large, use_typeS = "rate")

(p1_u1_varying_large | p2_u1_varying_large) / (p3_u1_varying_large | p4_u1_varying_large) + plot_layout(guides = "collect") & theme(legend.position = 'bottom')
```







```{r correlation between individual treatment effects and county random effects exists}

# function to generate correlated data

# 2×2: one treated state (s=1), one control (s=0)
simulate_data_2x2_one_group_predictor_correlated <- function(
  rho = 0.5, N_per_state = 40, K = 1,
  A0 = 0.0, A1 = 0.4,
  Xbar = matrix(c(-0.5, 0.8), nrow = 2, byrow = TRUE),
  u_1 = 1, u_0 = 0,
  sigma_c_treat = 1.0, sigma_c_ctrl = 0.6,
  lambda = 0.5, mu_delta = 3.0, sigma_delta = 1.6,
  sigma_y = 1.0, seed = 1L
){
  set.seed(seed)

  # Basic layout: S=2 states (1 treated, 2 control)
  S  <- 2L
  J_t <- N_per_state
  J_c <- N_per_state
  J   <- J_t + J_c

  is_treated       <- c(rep(1L, J_t), rep(0L, J_c))
  state_of_county  <- c(rep(1L, J_t), rep(2L, J_c))
  state_is_treated <- c(1L, 0L)

  # State-level effects u_s
  u <- c(u_1, u_0)  # NOTE: by your earlier code u <- c(u_0, u_1); choose your convention.
                    # Here I keep s=1 as treated (u_1), s=2 as control (u_0).
                    # If you prefer the other order, swap to c(u_0, u_1).

  # Allocate
  c     <- numeric(J)
  delta <- numeric(J)

  # ---- Joint draws for treated counties: (c_j, delta_j) ~ N( (0, mu_delta), Sigma ) ----
  # Sigma = [ sigma_c_treat^2, rho*sigma_c_treat*sigma_delta; rho*sigma_c_treat*sigma_delta, sigma_delta^2 ]
  if (!is.finite(rho) || abs(rho) >= 1) stop("rho must be finite with |rho| < 1.")
  if (sigma_c_treat <= 0 || sigma_delta <= 0) stop("sigma_c_treat and sigma_delta must be > 0.")
  if (sigma_c_ctrl  <= 0) stop("sigma_c_ctrl must be > 0.")

  Sigma_t <- matrix(c(
    sigma_c_treat^2,            rho * sigma_c_treat * sigma_delta,
    rho * sigma_c_treat * sigma_delta, sigma_delta^2
  ), nrow = 2, byrow = TRUE)

  # Numerical guard: if Sigma is nearly singular due to rho ~ +/-1, nudge
  eigvals <- eigen(Sigma_t, symmetric = TRUE, only.values = TRUE)$values
  if (min(eigvals) <= 0) {
    eps <- 1e-8
    # shrink off-diagonal slightly towards boundary to ensure PD
    off <- sign(rho) * min(abs(rho * sigma_c_treat * sigma_delta), sigma_c_treat * sigma_delta - eps)
    Sigma_t[1, 2] <- Sigma_t[2, 1] <- off
  }

  # Cholesky (upper) and simulate
  R <- chol(Sigma_t) # Upper triangular such that t(R) %*% R = Sigma
  Z <- matrix(rnorm(J_t * 2), ncol = 2)  # iid N(0,1)
  draws_treated <- Z %*% t(R)            # each row ~ N(0, Sigma)
  draws_treated[, 2] <- draws_treated[, 2] + mu_delta  # shift delta mean

  # Assign to treated block
  idx_t <- which(is_treated == 1L)
  c[idx_t]     <- draws_treated[, 1]
  delta[idx_t] <- draws_treated[, 2]

  # ---- Controls: c_j ~ N(0, sigma_c_ctrl^2), delta_j = 0 ----
  idx_c <- which(is_treated == 0L)
  c[idx_c]     <- rnorm(J_c, 0, sigma_c_ctrl)
  delta[idx_c] <- 0

  # ---- Build panel (two periods) ----
  county <- rep(seq_len(J), each = 2)
  post   <- rep(c(0L, 1L), times = J)
  D      <- is_treated[county] * post
  s_row  <- state_of_county[county]

  # Handle A1 as scalar or length-K vector
  A1_vec <- if (length(A1) == 1L) rep(A1, K) else as.numeric(A1)

  # Linear predictor from group-level predictors
  lin_x <- rowSums(
    Xbar[s_row, , drop = FALSE] *
      matrix(A1_vec, nrow = length(s_row), ncol = K, byrow = TRUE)
  )

  # Outcome
  mu <- A0 + lin_x + u[s_row] + c[county] + lambda * post + delta[county] * D
  y  <- rnorm(length(mu), mu, sigma_y)

  list(
    stan_data = list(
      N = length(y), J = J, S = 2L, K = K, y = y,
      county = as.integer(county), post = as.integer(post), D = as.integer(D),
      state_of_county = as.integer(state_of_county),
      Xbar = Xbar,
      is_treated = as.integer(is_treated),
      state_is_treated = as.integer(state_is_treated),
      J_treated = J_t,
      treated_idx = as.integer(idx_t),
      center_c = 0L,
      pool_u_within_group = 0L,
      big_sd_u = 10
    ),
    truth = list(
      A0 = A0, A1 = A1_vec, Xbar = Xbar, lambda = lambda, sigma_y = sigma_y,
      u = u, c = c, is_treated = is_treated, delta = delta,
      mu_delta = mu_delta, sigma_delta = sigma_delta, rho = rho,
      sigma_c_treat = sigma_c_treat, sigma_c_ctrl = sigma_c_ctrl
    )
  )
}
```




```{r quick compare with dependence}
quick_compare_with_correlation <- function(
  rho = 0.5, R = 50, N_per_state = 40, K = 1,
  A0 = 0.0, A1 = 0.4,
  Xbar = matrix(c(-0.5, 0.8), nrow = 2, byrow = TRUE),
  u_1 = 1, u_0 = 0,
  sigma_c_treat = 1.0, sigma_c_ctrl = 0.7,
  lambda = 2, mu_delta = 1.0, sigma_delta = 0.6,
  sigma_y = 1.0, seed0 = 1234,
  iter_warmup = 500, iter_sampling = 500, chains = 2
){
  ATT_rows   <- vector("list", R)
  DELTA_rows <- vector("list", R)

  for (r in seq_len(R)) {
    sim <- simulate_data_2x2_one_group_predictor_correlated(
      rho = rho, N_per_state = N_per_state, K = K,
      A0 = A0, A1 = A1, Xbar = Xbar,
      u_1 = u_1, u_0 = u_0,
      sigma_c_treat = sigma_c_treat, sigma_c_ctrl = sigma_c_ctrl,
      lambda = lambda, mu_delta = mu_delta, sigma_delta = sigma_delta,
      sigma_y = sigma_y, seed = seed0 + r
    )

    # unpack
    y      <- sim$stan_data$y               # length 2J
    county <- sim$stan_data$county          # length 2J
    post   <- sim$stan_data$post            # length 2J
    soc    <- sim$stan_data$state_of_county # length J (per county)
    S      <- sim$stan_data$S
    Nps    <- N_per_state
    treated_ids <- which(sim$truth$is_treated == 1L)

    # per-row state (1=treated state, 0=control) for helpers
    state_row <- soc[county]                  # length 2J, in {1,2}
    state01   <- as.integer(state_row == 1L)  # length 2J, in {0,1}

    # TWFE regression on aligned data (optional; not used for sigma_y here)
    df_reg <- data.frame(
      y = y,
      county = factor(county),
      post   = factor(post),
      state  = factor(state_row)
    )
    twfe_model <- lm(y ~ county + post + state:post, data = df_reg)

    # Use known simulation SD (homoskedastic)
    sigy <- sim$truth$sigma_y

    # --- Classical ATT (your signature) ---
    cls <- att_classical(
      y      = y,
      county = county,
      post   = post,
      state  = state01,      # IMPORTANT: per-row 0/1 indicator
      sigma_y = sigy,
      N_per_state = Nps
    )
    # cls is tibble {method, est, lo, hi}

    # --- Per-county DiD (your signature) ---
    dj <- per_county_did(
      y = y,
      county = county,
      post = post,
      state = state01,       # same 0/1 per-row state
      sigma_y = sigy,
      N_per_state = Nps,
      treated_ids = treated_ids
    )

    # --- Bayesian fit ---
    fitb <- fit_bayes_once(
      stan_data       = sim$stan_data,
      treated_idx     = sim$stan_data$treated_idx,
      iter_warmup     = iter_warmup,
      iter_sampling   = iter_sampling,
      chains          = chains,
      parallel_chains = chains,
      seed            = seed0 + 10000 + r,
      refresh         = 0
    )

    # ATT metrics
    true_ATT <- mean(sim$truth$delta[treated_ids])
    att_all <- dplyr::bind_rows(
      cls,
      tibble::tibble(method = "Bayes DiD", est = fitb$ATT$mean, lo = fitb$ATT$lo, hi = fitb$ATT$hi)
    ) |>
      dplyr::mutate(true = true_ATT,
                    bias = est - true,
                    squared_error = (est - true)^2)

    ATT_rows[[r]] <- att_all

    # δ_j: Bayes vs per-county DiD
    delta_bayes <- fitb$delta |>
      dplyr::filter(j %in% treated_ids) |>
      dplyr::rename(est_bayes = est, lo_bayes = lo, hi_bayes = hi)

    delta_did <- dj |>
      dplyr::rename(j = county, est_did = est, lo_did = lo, hi_did = hi)

    d_join <- dplyr::inner_join(delta_bayes, delta_did, by = "j") |>
      dplyr::mutate(
        truth = sim$truth$delta[j],
        sig_bayes = (0 > hi_bayes) | (0 < lo_bayes),
        sig_did   = (0 > hi_did)   | (0 < lo_did),
        typeS_bayes = sig_bayes & (sign(est_bayes) != sign(truth)),
        typeS_did   = sig_did   & (sign(est_did)   != sign(truth)),
        sq_bayes = (est_bayes - truth)^2,
        sq_did   = (est_did   - truth)^2
      )

    DELTA_rows[[r]] <- tibble::tibble(
      n_sig_bayes = sum(d_join$sig_bayes),
      n_sig_did   = sum(d_join$sig_did),
      typeS_bayes_count = sum(d_join$typeS_bayes),
      typeS_did_count   = sum(d_join$typeS_did),
      typeS_bayes_rate  = ifelse(sum(d_join$sig_bayes) > 0,
                                 sum(d_join$typeS_bayes) / sum(d_join$sig_bayes), NA_real_),
      typeS_did_rate    = ifelse(sum(d_join$sig_did) > 0,
                                 sum(d_join$typeS_did) / sum(d_join$sig_did), NA_real_),
      mse_bayes  = mean(d_join$sq_bayes),
      mse_did    = mean(d_join$sq_did),
      rmse_bayes = sqrt(mean(d_join$sq_bayes)),
      rmse_did   = sqrt(mean(d_join$sq_did)),
      rank_bayes = suppressWarnings(cor(d_join$est_bayes, d_join$truth, method = "spearman")),
      rank_did   = suppressWarnings(cor(d_join$est_did,   d_join$truth, method = "spearman"))
    )
  }

  ATT_all   <- dplyr::bind_rows(ATT_rows)
  DELTA_all <- dplyr::bind_rows(DELTA_rows)

  att_sum <- ATT_all |>
    dplyr::group_by(method) |>
    dplyr::summarize(
      rmse = sqrt(mean(squared_error)),
      mse  = mean(squared_error),
      bias = mean(bias),
      .groups = "drop"
    )

  delta_sum <- DELTA_all |>
    dplyr::summarize(
      n_sig_bayes_mean = mean(n_sig_bayes),
      typeS_bayes_count_mean = mean(typeS_bayes_count),
      typeS_bayes_rate_mean  = mean(typeS_bayes_rate, na.rm = TRUE),
      mse_bayes = mean(mse_bayes), rmse_bayes = mean(rmse_bayes),
      rank_bayes = mean(rank_bayes, na.rm = TRUE),
      n_sig_did_mean = mean(n_sig_did),
      typeS_did_count_mean = mean(typeS_did_count),
      typeS_did_rate_mean  = mean(typeS_did_rate, na.rm = TRUE),
      mse_did = mean(mse_did), rmse_did = mean(rmse_did),
      rank_did = mean(rank_did, na.rm = TRUE)
    )

  list(ATT = att_sum,
       DELTA = delta_sum,
       ATT_raw = ATT_all,
       DELTA_raw = DELTA_all)
}



```




```{r run grid with correlation}

library(purrr); library(dplyr); library(tidyr);
run_grid_with_correlation <- function(rho, lambdas, u1_grid, R = 50, N_per_state = 40, ...) {
  purrr::map2_dfr(lambdas, u1_grid, function(lam, u1) {
    res <- quick_compare_with_correlation(
      rho = rho, R = R, N_per_state = N_per_state,
      lambda = lam, u_1 = u1, ...
    )
    tibble::tibble(lambda = lam, u_1 = u1) |>
      dplyr::bind_cols(res$DELTA)
  })
}
```



```{r run experiments}
# generate data with rho to be 0.5


# fit model 1 as we did before. Independent priors. Do not model the correlation between the individual treatment effects and random effects.

rho <- 0.5
start_time <- Sys.time()
lambdas_large <- seq(-30, 30, by = 10)
u1_grid_large <- rep(1, length(lambdas_large))
grid_out <- run_grid_with_correlation(
  rho = rho,
  lambdas = lambdas_large,
  u1_grid = u1_grid_large,
  R = 5,
  iter_warmup = 800, iter_sampling = 600, chains = 2
)
end_time <- Sys.time()
end_time - start_time

```

```{r plot the results with rho to be 0.5}

saveRDS(grid_out, file = "data/grid_out_lambda_varying_with_correlation.rds")

p1_lambda_varying_with_correlation <- plot_delta_mean_squared_error(grid_out, x_axis = "lambda")

p2_lambda_varying_with_correlation <- plot_delta_rank_correlation(grid_out, x_axis = "lambda")

p3_lambda_varying_with_correlation <- plot_delta_significant_counts(grid_out, x_axis = "lambda")

p4_lambda_varying_with_correlation <- plot_delta_type_S_errors(grid_out, x_axis = "lambda" , use_typeS = "rate")

(p1_lambda_varying_with_correlation | p2_lambda_varying_with_correlation) / (p3_lambda_varying_with_correlation | p4_lambda_varying_with_correlation) + plot_layout(guides = "collect") & theme(legend.position = 'bottom')

ggsave("images/grid_out_lambda_varying_with_correlation.png", width = 10, height = 8)

```


```{r run a large-scale experiments with rho to be 0.5}

rho <- 0.5
start_time <- Sys.time()
lambdas_large <- seq(-40, 40, by = 10)

# u1_grid_large is simply repetition of ones.
u1_grid_large <- rep(1, length(lambdas_large))


grid_out <- run_grid_with_correlation(
  rho = rho,
  lambdas = lambdas_large,
  u1_grid = u1_grid_large,
  R = 10,
  iter_warmup = 800, iter_sampling = 600, chains = 2
)
end_time <- Sys.time()
end_time - start_time

```

```{r plot the results with rho to be 0.5}

saveRDS(grid_out, file = "data/grid_out_lambda_varying_with_correlation.rds")

p1_lambda_varying_with_correlation <- plot_delta_mean_squared_error(grid_out, x_axis = "lambda")

p2_lambda_varying_with_correlation <- plot_delta_rank_correlation(grid_out, x_axis = "lambda")

p3_lambda_varying_with_correlation <- plot_delta_significant_counts(grid_out, x_axis = "lambda")

p4_lambda_varying_with_correlation <- plot_delta_type_S_errors(grid_out, x_axis = "lambda" , use_typeS = "rate")

(p1_lambda_varying_with_correlation | p2_lambda_varying_with_correlation) / (p3_lambda_varying_with_correlation | p4_lambda_varying_with_correlation) + plot_layout(guides = "collect") & theme(legend.position = 'bottom')

ggsave("images/grid_out_lambda_varying_with_correlation.png", width = 10, height = 8)

```


```{r run the experiments and plot the results with different rhos}
library(dplyr)
library(tidyr)
library(purrr)
library(ggplot2)

# ------------------------------------------------------------
# Runner: fixed u_1, vary rho and lambda
# ------------------------------------------------------------
run_rho_lambda_grid <- function(
  rhos,
  lambdas,
  u1_fixed,
  R = 50, N_per_state = 40, K = 1,
  A0 = 0.0, A1 = 0.4,
  Xbar = matrix(c(-0.5, 0.8), nrow = 2, byrow = TRUE),
  u_0 = 0,
  sigma_c_treat = 1.0, sigma_c_ctrl = 0.7,
  mu_delta = 1.0, sigma_delta = 0.6,
  sigma_y = 1.0, seed0 = 1234,
  iter_warmup = 500, iter_sampling = 500, chains = 2
){
  grid <- tidyr::expand_grid(rho = rhos, lambda = lambdas)

  purrr::pmap_dfr(
    grid,
    function(rho, lambda){
      out <- quick_compare_with_correlation(
        rho = rho, R = R, N_per_state = N_per_state, K = K,
        A0 = A0, A1 = A1, Xbar = Xbar,
        u_1 = u1_fixed, u_0 = u_0,
        sigma_c_treat = sigma_c_treat, sigma_c_ctrl = sigma_c_ctrl,
        lambda = lambda, mu_delta = mu_delta, sigma_delta = sigma_delta,
        sigma_y = sigma_y, seed0 = seed0,
        iter_warmup = iter_warmup, iter_sampling = iter_sampling, chains = chains
      )
      out$DELTA |>
        mutate(rho = rho, lambda = lambda, u_1 = u1_fixed)
    }
  )
}
```


```{r}
# ------------------------------------------------------------
# Plotting helpers (facet by lambda; u1 is fixed)
# Assumes quick_compare_with_correlation() produced these columns:
#   rmse_b (BDiD), rmse_f (DiD),
#   typeS_bayes_did_rate_mean, typeS_did_rate_mean,
#   n_sig_bayes_did_mean, n_sig_did_mean,
#   rank_b (BDiD), rank_f (DiD)
# ------------------------------------------------------------

# ---- 0) Column harmonizer (as before) ----
harmonize_delta_cols_model_1 <- function(df) {
  rn <- function(old, new) if (all(old %in% names(df)) && !(new %in% names(df)))
    df <<- dplyr::rename(df, !!new := dplyr::all_of(old))
  rn("mse_bayes",  "mse_b");   rn("rmse_bayes","rmse_b"); rn("rank_bayes","rank_b")
  rn("n_sig_bayes_mean","n_sig_bayes_did_mean")
  rn("typeS_bayes_count_mean","typeS_bayes_did_count_mean")
  rn("typeS_bayes_rate_mean","typeS_bayes_did_rate_mean")
  rn("mse_did","mse_f"); rn("rmse_did","rmse_f"); rn("rank_did","rank_f")
  if (!"rmse_b"%in%names(df) && "mse_b"%in%names(df)) df$rmse_b <- sqrt(df$mse_b)
  if (!"rmse_f"%in%names(df) && "mse_f"%in%names(df)) df$rmse_f <- sqrt(df$mse_f)
  df
}

# ---- 1) Internal: common scaffolding for rho–λ plots ----
.make_panel <- function(df, ycols, ylab, point_size, linewidth, rhos = NULL) {
  df <- harmonize_delta_cols_model_1(df)
  if (!is.null(rhos)) df <- dplyr::filter(df, rho %in% rhos)

  # long format with method labels
  long <- df |>
    dplyr::select(rho, lambda, dplyr::all_of(ycols)) |>
    tidyr::pivot_longer(cols = dplyr::all_of(ycols),
                        names_to = "method", values_to = "value") |>
    dplyr::mutate(method = dplyr::recode(method,
      "rmse_b"="BDiD","rmse_f"="DiD",
      "mse_b"="BDiD","mse_f"="DiD",
      "rank_b"="BDiD","rank_f"="DiD",
      "n_sig_bayes_did_mean"="BDiD","n_sig_did_mean"="DiD",
      "typeS_bayes_did_rate_mean"="BDiD","typeS_did_rate_mean"="DiD",
      "typeS_bayes_did_count_mean"="BDiD","typeS_did_count_mean"="DiD"
    ))

  # global y-limits across all facets to lock scales
  y_min <- min(long$value, na.rm = TRUE)
  y_max <- max(long$value, na.rm = TRUE)
  # small padding
  pad <- 0.04 * (y_max - y_min + 1e-9)
  y_lim <- c(y_min - pad, y_max + pad)

  ggplot(long, aes(x = lambda, y = value, linetype = method, shape = method)) +
    geom_line(linewidth = linewidth) +
    geom_point(size = point_size) +
    scale_linetype_manual(values = c(BDiD = "dashed", DiD = "solid")) +
    scale_shape_manual(values = c(BDiD = 17, DiD = 16)) +
    labs(x = expression(lambda), y = ylab, linetype = "Method", shape = "Method") +
    coord_cartesian(ylim = y_lim) +              # FIXED global y-limit across facets
    facet_wrap(~ rho, scales = "fixed") +        # one subplot per ρ
    theme_minimal(base_size = 12) +
    theme(legend.position = "bottom")
}

# ---- 2) RMSE(δ_j) vs λ, faceted by ρ ----
plot_rmse_vs_rho <- function(df, point_size, linewidth, rhos = NULL) {
  .make_panel(df, ycols = c("rmse_b","rmse_f"), ylab = "RMSE",  point_size, linewidth, rhos = rhos)
}

# ---- 3) Rank correlation(δ̂_j, δ_j) vs λ, faceted by ρ ----
plot_rank_vs_rho <- function(df, point_size, linewidth, rhos = NULL) {
  .make_panel(df, ycols = c("rank_b","rank_f"), ylab = "Spearman rank corr.", point_size = point_size, linewidth = linewidth, rhos = rhos)
}

# ---- 4) # significant δ_j vs λ, faceted by ρ ----
plot_nsig_vs_rho <- function(df,  point_size, linewidth, rhos = NULL) {
  .make_panel(df, ycols = c("n_sig_bayes_did_mean","n_sig_did_mean"),
              ylab = "Mean # significant", point_size = point_size, linewidth = linewidth, rhos = rhos)
}

# ---- 5) Type-S errors vs λ, faceted by ρ (choose "rate" or "count") ----
plot_typeS_vs_rho <- function(df, point_size, linewidth, rhos = NULL, which = c("rate","count")) {
  which <- match.arg(which)
  if (which == "rate") {
    .make_panel(df,
      ycols = c("typeS_bayes_did_rate_mean","typeS_did_rate_mean"),
      ylab  = "Type-S error rate",  
      point_size = point_size, 
      linewidth = linewidth,
      rhos  = rhos
    )
  } else {
    .make_panel(df,
      ycols = c("typeS_bayes_did_count_mean","typeS_did_count_mean"),
      ylab  = "Mean Type-S error (count)",  
      point_size = point_size, 
      linewidth = linewidth,
      rhos  = rhos
    )
  }
}
```



```{r}
# ------------------------------------------------------------
# Example
# ------------------------------------------------------------
set.seed(2026)
rhos    <- c(-0.99, -0.9, -0.5, -0.05, -0.01,  0, 0.01,  0.05, 0.5, 0.95, 0.99)
lambdas <- seq(-100, 100, by = 10)
u1_fixed <- 1
#

start_time <- Sys.time()
df_gl <- run_rho_lambda_grid(
   rhos = rhos, lambdas = lambdas, u1_fixed = u1_fixed,
   R = 30, N_per_state = 40,
   iter_warmup = 400, iter_sampling = 400, chains = 2
 )

end_time <- Sys.time()
time_1 <- end_time - start_time
print(time_1)
```
```{r}
library(patchwork)

df_gl <- harmonize_delta_cols_model_1(df_gl)

# Save the harmonized data
saveRDS(df_gl, file = file.path("data", "df_gl.rds"))
readr::write_csv(df_gl, file = file.path("data", "df_gl.csv"))

```


```{r}

df_gl <- harmonize_delta_cols_model_1(df_gl)
p_rmse  <- plot_rmse_vs_rho(df = df_gl, point_size = 1.5, linewidth = 0.7,  rhos =   c(-0.99, -0.9, -0.5, -0.05, -0.01,  0, 0.01,  0.05, 0.5, 0.95, 0.99) )
p_rank  <- plot_rank_vs_rho(df = df_gl, point_size = 1.5, linewidth = 0.7,  rhos =  c(-0.99, -0.9, -0.5, -0.05, -0.01,  0, 0.01,  0.05, 0.5, 0.95, 0.99))
p_nsig  <- plot_nsig_vs_rho(df = df_gl, point_size = 1.5, linewidth = 0.7,  rhos =   c(-0.99, -0.9, -0.5, -0.05, -0.01,  0, 0.01,  0.05, 0.5, 0.95, 0.99))
p_typeS <- plot_typeS_vs_rho(df = df_gl, point_size = 1.5, linewidth = 0.7, rhos =   c(-0.99, -0.9, -0.5, -0.05, -0.01,  0, 0.01,  0.05, 0.5, 0.95, 0.99), which = "rate")



# ---- Harmonize, subset by rhos of interest, and save the data ----

rhos_to_plot <- c(-0.99, -0.9, -0.5, -0.05, -0.01,  0, 0.01,  0.05, 0.5, 0.95, 0.99)

# Create output folder
outdir <- "figs_rho_lambda_model_1"
if (!dir.exists(outdir)) dir.create(outdir, recursive = TRUE)
```



```{r}

df_gl <- readRDS("data/df_gl.rds")

```

```{r}
# ---- Save each separately (PNG + PDF) ----
ggsave(filename = file.path("images/model1", "rmse_vs_lambda_by_rho.png"),
       plot = p_rmse, width = 8, height = 6, dpi = 300, bg = "white")
ggsave(filename = file.path("images/model1", "rmse_vs_lambda_by_rho.pdf"),
       plot = p_rmse, width = 8, height = 6, device = cairo_pdf, bg = "white")

ggsave(filename = file.path("images/model1", "rankcorr_vs_lambda_by_rho.png"),
       plot = p_rank, width = 8, height = 6, dpi = 300, bg = "white")
ggsave(filename = file.path("images/model1", "rankcorr_vs_lambda_by_rho.pdf"),
       plot = p_rank, width = 8, height = 6, device = cairo_pdf, bg = "white")

ggsave(filename = file.path("images/model1", "nsignificant_vs_lambda_by_rho.png"),
       plot = p_nsig, width = 8, height = 6, dpi = 300, bg = "white")
ggsave(filename = file.path("images/model1", "nsignificant_vs_lambda_by_rho.pdf"),
       plot = p_nsig, width = 8, height = 6, device = cairo_pdf, bg = "white")

ggsave(filename = file.path("images/model1", "typeSrate_vs_lambda_by_rho.png"),
       plot = p_typeS, width = 8, height = 6, dpi = 300, bg = "white")
ggsave(filename = file.path("images/model1", "typeSrate_vs_lambda_by_rho.pdf"),
       plot = p_typeS, width = 8, height = 6, device = cairo_pdf, bg = "white")

p_rmse  
p_rank  
p_nsig  
p_typeS 

```

```{r}

data_correlated <- simulate_data_2x2_one_group_predictor_correlated(rho = -0.99)

cor(data_correlated$truth$c, data_correlated$truth$delta)

```
```{r}
corr <- 0.2

n_t <- 10 
c_i <- rnorm(n_t)


df$treated <- rep(ifelse(corr * scale(c_i) + sqrt(1 - corr^2) * rnorm(n_t) > 0, 1, 0), each = n_t)
	
```








